---
title: "Assay distributions"
author: "Kara Tsang, Kat Holt"
date: "Last compiled on `r format(Sys.time(), '%d/%m/%y')`"
output: 
  html_document: 
    df_print: paged
    fig_width: 9
    fig_height: 8
    mode: selfcontained
    highlight: pygments
    theme: 
      bootswatch: flatly
      code_font:
        google: "JetBrains Mono"
editor_options: 
  chunk_output_type: inline
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 1600px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = 'png', echo = TRUE, message = FALSE, warning = FALSE,
                      cache = FALSE, fig.align = 'center')
knitr::opts_knit$set(root.dir = getwd())

# Data wrangling
library(tidyverse)
library(dplyr)
library(patchwork)
```

# load data
``` {r}
geno <- read_delim("tables/TableS2_genotypes.tsv")
pheno <- read_delim("tables/TableS3_phenotypes_metadata.tsv")
cipro_antibiogram <- full_join(geno,pheno)

```

## colours
``` {r}

c25 <- c(
  "dodgerblue2", "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)
```

# Relabel MIC datasets
```{r}
cipro_antibiogram_MIC <- cipro_antibiogram

cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "KASPAH" ] <- "KASPAH (V=1172)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "KLEBGAP" ] <- "KLEBGAP (V=92, E=4)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "India_GHRU" ] <- "India_GHRU (V=831)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Leangapichart_2021" ] <- "Leangapichart_2021 (V=38)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Philippines" ] <- "Philippines (V=227)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Torok_2020" ] <- "Torok_2020 (V=1296)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Egli" ] <- "Egli (V=127)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Long" ] <- "Long (P=1595)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Denmark_MedVetKlebs" ] <- "Denmark_MedVetKlebs (S=49)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Mathers" ] <- "Mathers (S=27)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Italy_MVK_MIC" ] <- "Italy_MVK_MIC (S=85)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "BSAC" ] <- "BSAC (A=247)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Milan_2019" ] <- "Milan_2019 (S=98)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Wright" ] <- "Wright (V=80)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Colombia_GHRU" ] <- "Colombia_GHRU (V=394)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "BARNARDS" ] <- "BARNARDS (A=251)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "SPARK" ] <- "SPARK (P=1956)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Ellington" ] <- "Ellington (V=70)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Nigeria_GHRU" ] <- "Nigeria_GHRU (V=108)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Stoesser_2013" ] <- "Stoesser_2013 (P=69)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Heinz_2019" ] <- "Heinz_2019 (V=178)"
cipro_antibiogram_MIC$index[cipro_antibiogram_MIC$index == "Bachman_2022" ] <- "Bachman_2022 (S=85)"



cipro_antibiogram_MIC$index <- factor(cipro_antibiogram_MIC$index, levels = c(
  "Torok_2020 (V=1296)",
  "KASPAH (V=1172)",
  "India_GHRU (V=831)",
  "Colombia_GHRU (V=394)",
  "Philippines (V=227)",
  "Heinz_2019 (V=178)",
  "Egli (V=127)",
  "Nigeria_GHRU (V=108)",
  "KLEBGAP (V=92, E=4)",
  "Wright (V=80)",
  "Ellington (V=70)",
  "Leangapichart_2021 (V=38)",
  "SPARK (P=1956)", 
  "Long (P=1595)",
  "Stoesser_2013 (P=69)",
  "Milan_2019 (S=98)",
  "Bachman_2022 (S=85)",
  "Italy_MVK_MIC (S=85)",
  "Denmark_MedVetKlebs (S=49)",
  "Mathers (S=27)",
  "BARNARDS (A=251)",
  "BSAC (A=247)"
 ))

```


# MIC
``` {r}
MIC_dist <- cipro_antibiogram_MIC %>%
  filter(Laboratory.Typing.Method != "Disk diffusion") %>%
  mutate(
    Measurement = replace(Measurement, Measurement == 0.032, 0.03),
    Measurement = replace(Measurement, Measurement == 0.064, 0.06),
    Measurement = replace(Measurement, Measurement == 0.12, 0.125),

    # Merge all 0.5 MIC values into one category
    Measurement.sign = ifelse(Measurement == 0.5, "=", Measurement.sign),

    MIC_label = ifelse(
      Measurement.sign %in% c("-", "=", "=="),
      as.character(Measurement),
      paste0(Measurement.sign, Measurement)
    ),

    order_value = case_when(
      Measurement.sign %in% c("<", "<=") ~ Measurement - 0.01,
      Measurement.sign %in% c(">", ">=") ~ Measurement + 0.01,
      TRUE ~ Measurement
    )
  ) %>%
  group_by(
    Laboratory.Typing.Platform,
    Measurement,
    Measurement.sign,
    MIC_label,
    order_value,
    index
  ) %>%
  count() %>%
  ungroup() %>%
  arrange(order_value) %>%
  mutate(
    MIC_label = factor(
      MIC_label,
      levels = unique(MIC_label)
    )
  ) %>%
  ggplot(aes(x = MIC_label, y = n, fill = index)) + 
  geom_bar(position = "stack", stat = "identity") +
  facet_wrap(~Laboratory.Typing.Platform, ncol = 1, scales = "free_y") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    x = "MIC (mg/L)", 
    y = "Number of genomes", 
    fill = "Dataset"
  ) +
  scale_fill_manual(values = c25[1:22]) +
  geom_vline(xintercept = 11.5, linetype = 2) +
  geom_vline(xintercept = 14.5, linetype = 2)


```


# Relabel DD datasets
```{r}
cipro_antibiogram_DD <- cipro_antibiogram

cipro_antibiogram_DD$index[cipro_antibiogram_DD$index == "Huynh_2020" ] <- "Huynh_2020 (N=262)"
cipro_antibiogram_DD$index[cipro_antibiogram_DD$index == "KLEBGAP" ] <- "KLEBGAP (N=2379)"
cipro_antibiogram_DD$index[cipro_antibiogram_DD$index == "Philippines_DD" ] <- "Philippines_DD (N=28)"
cipro_antibiogram_DD$index[cipro_antibiogram_DD$index == "Potter_2018" ] <- "Potter_2018 (N=52)"
cipro_antibiogram_DD$index[cipro_antibiogram_DD$index == "perdigao_2022" ] <- "perdigao_2022 (N=334)"
cipro_antibiogram_DD$index[cipro_antibiogram_DD$index == "sands2021_arthropods_DD" ] <- "sands2021_arthropods_DD (N=18)"

cipro_antibiogram_DD$index <- factor(cipro_antibiogram_DD$index, levels = c(
  "KLEBGAP (N=2379)",
  "perdigao_2022 (N=334)",
  "Huynh_2020 (N=262)",
  "Potter_2018 (N=52)",
  "Philippines_DD (N=28)",
  "sands2021_arthropods_DD (N=18)"

 ))

```


# Disk diffusion
``` {r}
DD_dist <- cipro_antibiogram_DD %>%
  filter(Laboratory.Typing.Method=="Disk diffusion") %>%
  group_by(Measurement, index) %>% count() %>%
  ggplot(aes(x=factor(Measurement), y=n, fill=index)) + 
  geom_bar(position="stack", stat="identity") +
  theme_bw() +
  labs(x="Disk diffusion zone diameter (mm)", y="Number of genomes", fill="Dataset") +
  scale_fill_manual(values=c25[10:15]) +
  geom_vline(xintercept=16.5, linetype=2) +
  #annotate("text",x=16, y=400, label="R") +
  geom_vline(xintercept=19.5, linetype=2) +
  #annotate("text",x=20, y=400, label="S") + 
  scale_x_discrete(breaks=seq(5,45,5))

```

# Figure S2 - MIC and disk distribution
``` {r figs2}

MIC_dist
DD_dist

(MIC_dist + ggtitle(label="a) Minimum inhibitory concentration distribution"))  / (DD_dist + ggtitle(label="b) Disk diffusion distribution"))


ggsave(width=9, height=10, file="figs/FigS2_MIC_DD_distribution.pdf")
ggsave(width=9, height=10, file="figs/FigS2_MIC_DD_distribution.png")

```


# Figure S3 - concordant and discordant plotting

# Discordant phenotype - MIC test vs MIC retest
```{r, discordant_mic_mic_1}

# Create discordant MIC dataset
cipro_antibiogram_MIC_discordant <- cipro_antibiogram_MIC %>%
  filter(Laboratory.Typing.Method != "Disk diffusion") %>%
  mutate(
    Measurement = replace(Measurement, Measurement == 0.032, 0.03),
    Measurement = replace(Measurement, Measurement == 0.064, 0.06),
    Measurement = replace(Measurement, Measurement == 0.12, 0.125),

    MIC_label = ifelse(
      Measurement.sign %in% c("-", "=", "=="),
      as.character(Measurement),
      paste0(Measurement.sign, Measurement)
    )
  ) %>%
  select(
    strain,
    MIC_label,
    DiscordantRetesting_OriginalMeasurement
  ) %>%
  filter(!is.na(DiscordantRetesting_OriginalMeasurement)) %>%
  rename(
    Test = DiscordantRetesting_OriginalMeasurement,
    `Re-test` = MIC_label
  )


# Convert to long format
cipro_antibiogram_MIC_discordant_long <- 
  cipro_antibiogram_MIC_discordant %>%
  pivot_longer(
    cols = c(Test, `Re-test`),
    names_to = "Method",
    values_to = "Measurement"
  )


# Get all MIC values from both Test and Re-test
mic_labels <- unique(
  c(
    cipro_antibiogram_MIC_discordant$Test,
    cipro_antibiogram_MIC_discordant$`Re-test`
  )
)


# Extract numeric portion for ordering
mic_order <- sapply(
  mic_labels,
  function(x) {
    
    value <- as.numeric(
      gsub("[^0-9.]", "", x)
    )
    
    if (grepl("^<", x)) {
      value - 0.001
    } else if (grepl("^>", x)) {
      value + 0.001
    } else {
      value
    }
  }
)


# Order MIC labels numerically
mic_levels <- mic_labels[order(mic_order)]


# Make MIC a categorical variable
cipro_antibiogram_MIC_discordant_long <- 
  cipro_antibiogram_MIC_discordant_long %>%
  mutate(
    Measurement = factor(
      Measurement,
      levels = mic_levels,
      ordered = TRUE
    ),
    
    Method = factor(
      Method,
      levels = c("Test", "Re-test")
    )
  )
# Aggregate to get count of strains per Test/Re-test combination
cipro_antibiogram_MIC_discordant_wide <- 
  cipro_antibiogram_MIC_discordant_long %>%
  pivot_wider(
    names_from = Method,
    values_from = Measurement
  ) %>%
  count(Test, `Re-test`, name = "n_strains")

# Helper: classify an MIC label based on clinical breakpoints
# Resistant: >=1 mg/L
# Intermediate: 0.5 mg/L
# Susceptible: <0.5 mg/L
classify_mic <- function(label) {

  numeric_value <- as.numeric(
    gsub("[^0-9.]", "", label)
  )

  ifelse(
    numeric_value >= 1,
    "R",
    ifelse(
      numeric_value == 0.5,
      "I",
      "S"
    )
  )
}


# Aggregate counts and add Test -> Re-test classification
cipro_antibiogram_MIC_discordant_wide <- 
  cipro_antibiogram_MIC_discordant_long %>%
  pivot_wider(
    names_from = Method,
    values_from = Measurement
  ) %>%
  count(Test, `Re-test`, name = "n_strains") %>%
  mutate(

    # Classify Test and Re-test independently
    Test_category   = classify_mic(as.character(Test)),
    Retest_category = classify_mic(as.character(`Re-test`)),

    # Label the direction of change
    Transition = ifelse(
      Test_category == Retest_category,
      "No change",
      paste(Test_category, "to", Retest_category)
    )
  )

scale_colour_manual(
  name = "Test → Re-test",
  values = c(
    "No change" = "#8EC476",
    "S to I"    = "#E5C07B",
    "S to R"    = "#D66D27",
    "I to S"    = "#E5C07B",
    "I to R"    = "#D66D27",
    "R to S"    = "#D66D27",
    "R to I"    = "#E5C07B"
  )
)

# Scatter plot: Test vs Re-test, sized by strain count,
# coloured by Test -> Re-test transition
discordant_mic_mic_1 <-
  ggplot(
    cipro_antibiogram_MIC_discordant_wide,
    aes(
      x = Test,
      y = `Re-test`,
      size = n_strains,
      colour = Transition
    )
  ) +
  geom_point(
    alpha = 0.7
  ) +
    scale_size_continuous(
    labels = scales::number_format(accuracy = 1) # Forces integers/absolute numbers
  )+
  geom_vline(
    xintercept = which(mic_levels == "1"),
    linetype = "dashed"
  ) +
  geom_hline(
    yintercept = which(mic_levels == "1"),
    linetype = "dashed"
  ) +
    geom_vline(
    xintercept = which(mic_levels == ">=0.25"),
    linetype = "dashed"
  ) +
  geom_hline(
    yintercept = which(mic_levels == ">=0.25"),
    linetype = "dashed"
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  scale_size_continuous(
    name = "Number of strains",
    range = c(2, 10),
    breaks = scales::pretty_breaks()
  ) +
scale_colour_manual(
  name = "Phenotype change (Test → Re-test)",
  values = c(
    "No change" = "#8EC476",
    "S to I"    = "#E5C07B",
    "S to R"    = "#D66D27",
    "I to S"    = "#E5C07B",
    "I to R"    = "#D66D27",
    "R to S"    = "#D66D27",
    "R to I"    = "#E5C07B"
  )
) +
  theme_classic() +
  labs(
    x = "Test MIC (mg/L)",
    y = "Re-test MIC (mg/L)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotates labels

discordant_mic_mic_1


```

# Discordant phenotype - Disk test vs disk retest
```{r, discordant_disk_disk_2}

# Create discordant DD dataset
cipro_antibiogram_DD_discordant <- cipro_antibiogram_DD %>%
  filter(Laboratory.Typing.Method == "Disk diffusion") %>%
  select(
    strain,
    Measurement,
    DiscordantRetesting_OriginalMeasurement
  ) %>%
  filter(!is.na(DiscordantRetesting_OriginalMeasurement)) %>%
  rename(
    Test = DiscordantRetesting_OriginalMeasurement,
    `Re-test` = Measurement
  ) %>%
  mutate(
    Test = as.numeric(Test),
    `Re-test` = as.numeric(`Re-test`)
  ) %>%
  # Remove MIC measurements
  filter(Test > 2)


# Convert to long format
cipro_antibiogram_DD_discordant_long <-
  cipro_antibiogram_DD_discordant %>%
  pivot_longer(
    cols = c(Test, `Re-test`),
    names_to = "Method",
    values_to = "Measurement"
  )


# Define the order of DD labels for both axes
DD_labels <- sort(unique(
  c(
    cipro_antibiogram_DD_discordant$Test,
    cipro_antibiogram_DD_discordant$`Re-test`
  )
))


# Convert to ordered factors, aggregate, and classify
cipro_antibiogram_DD_discordant_wide <-
  cipro_antibiogram_DD_discordant_long %>%
  mutate(
    Measurement = factor(
      Measurement,
      levels = DD_labels,
      ordered = TRUE
    ),
    Method = factor(
      Method,
      levels = c("Test", "Re-test")
    )
  ) %>%
  pivot_wider(
    names_from = Method,
    values_from = Measurement
  ) %>%
  count(Test, `Re-test`, name = "n_strains") %>%
  mutate(

    # Classify Test and Re-test based on DD breakpoints
    # Resistant: <=21 mm
    # Intermediate: 22-24 mm
    # Susceptible: >=25 mm

    Test_category = case_when(
      as.numeric(as.character(Test)) <= 21 ~ "R",
      as.numeric(as.character(Test)) >= 22 &
        as.numeric(as.character(Test)) <= 24 ~ "I",
      as.numeric(as.character(Test)) >= 25 ~ "S"
    ),

    Retest_category = case_when(
      as.numeric(as.character(`Re-test`)) <= 21 ~ "R",
      as.numeric(as.character(`Re-test`)) >= 22 &
        as.numeric(as.character(`Re-test`)) <= 24 ~ "I",
      as.numeric(as.character(`Re-test`)) >= 25 ~ "S"
    ),

    # Label unchanged results as "No change"
    # and changed results by direction
    Transition = ifelse(
      Test_category == Retest_category,
      "No change",
      paste(Test_category, "to", Retest_category)
    )
  )


# Position of the 21 mm threshold on the discrete axes
threshold_position <- which(DD_labels == 21)

size_values <- sort(
  unique(cipro_antibiogram_DD_discordant_wide$n_strains)
)

size_mapping <- setNames(
  seq(
    2,
    10,
    length.out = length(size_values)
  ),
  size_values
)


# Scatter plot: Test vs Re-test,
# sized by strain count,
# coloured by Test -> Re-test transition
discordant_disk_disk_2 <-
  ggplot(
    cipro_antibiogram_DD_discordant_wide,
    aes(
      x = Test,
      y = `Re-test`,
      size = factor(n_strains),
      colour = Transition
    )
  ) +
  
  geom_point(alpha = 0.7) +

  # Resistant / Intermediate boundary
  geom_vline(
    xintercept = threshold_position + 0.5,
    linetype = "dashed"
  ) +
  geom_hline(
    yintercept = threshold_position + 0.5,
    linetype = "dashed"
  ) +

  # Intermediate / Susceptible boundary (24 -> 25)
  geom_vline(
    xintercept = which(DD_labels == 24) + 0.5,
    linetype = "dashed"
  ) +
  geom_hline(
    yintercept = which(DD_labels == 24) + 0.5,
    linetype = "dashed"
  ) +

  scale_x_discrete(
    drop = FALSE
  ) +
  
  scale_y_discrete(
    drop = FALSE
  ) +

  # Discrete size scale:
  # one unique circle size for each actual strain count
  scale_size_manual(
    name = "Number of strains",
    values = size_mapping,
    drop = FALSE
  ) +

  scale_colour_manual(
    name = "Test → Re-test",
    values = c(
      "No change" = "#8EC476",
      "R to I"    = "#E5C07B",
      "I to R"    = "#E5C07B",
      "I to S"    = "#E5C07B",
      "S to I"    = "#E5C07B",
      "R to S"    = "#D66D27",
      "S to R"    = "#D66D27"
    )
  ) +

  theme_classic() +

  labs(
    x = "Test disk diffusion (mm)",
    y = "Re-test disk diffusion (mm)"
  )


discordant_disk_disk_2
```

# Discordant phenotype - MIC test vs disk retest
```{r, eval=FALSE, discordant_mic_disk_3}
# Create discordant MIC/disk diffusion dataset
cipro_antibiogram_DD_discordant <- cipro_antibiogram_DD %>%
  filter(Laboratory.Typing.Method == "Disk diffusion") %>%
  select(
    strain,
    Measurement,
    DiscordantRetesting_OriginalMeasurement
  ) %>%
  filter(!is.na(DiscordantRetesting_OriginalMeasurement)) %>%
  rename(
    Test = DiscordantRetesting_OriginalMeasurement,
    `Re-test` = Measurement
  ) %>%
  mutate(
    Test = as.numeric(Test),
    `Re-test` = as.numeric(`Re-test`)
  ) %>%
  # Keep only MIC measurements for Test (<2 mg/L)
  filter(Test < 2)

# Build ordered factor levels numerically for both axes
test_levels  <- sort(unique(cipro_antibiogram_DD_discordant$Test))
retest_levels <- sort(unique(cipro_antibiogram_DD_discordant$`Re-test`))

cipro_antibiogram_DD_discordant <- cipro_antibiogram_DD_discordant %>%
  mutate(
    Test = factor(Test, levels = test_levels, ordered = TRUE),
    `Re-test` = factor(`Re-test`, levels = retest_levels, ordered = TRUE)
  )

# Aggregate counts directly — no pivot_longer needed
cipro_antibiogram_DD_discordant_wide <- 
  cipro_antibiogram_DD_discordant %>%
  count(Test, `Re-test`, name = "n_strains")

# MIC Test classification # Resistant: >=1 mg/L # Intermediate: 0.5 mg/L # Susceptible: <0.5 mg/L
# Disk diffusion Re-test classification # Resistant: <=21 mm # Intermediate: 22-24 mm # Susceptible: >=25 mm 
# Label unchanged results as "No change" # and changed results by direction 
cipro_antibiogram_DD_discordant_wide <- 
  cipro_antibiogram_DD_discordant_wide %>% mutate(
  Test_category = case_when( as.numeric(as.character(Test)) >= 1 ~ "R", as.numeric(as.character(Test)) == 0.5 ~ "I", as.numeric(as.character(Test)) < 0.5 ~ "S" ),
  Retest_category = case_when( as.numeric(as.character(`Re-test`)) <= 21 ~ "R", as.numeric(as.character(`Re-test`)) >= 22 & as.numeric(as.character(`Re-test`)) <= 24 ~ "I", as.numeric(as.character(`Re-test`)) >= 25 ~ "S" ),
  Transition = ifelse( Test_category == Retest_category, "No change", paste(Test_category, "to", Retest_category) ) )
  
  

# Scatter plot: Test MIC vs Re-test disk diffusion

discordant_mic_disk_3 <- ggplot( cipro_antibiogram_DD_discordant_wide, aes( x = Test, y = `Re-test`, size = n_strains, colour = Transition ) ) + geom_point(alpha = 0.7) +
    scale_size_continuous(
    labels = scales::number_format(accuracy = 1) # Forces integers/absolute numbers
  )+


geom_vline( xintercept = which(test_levels == 0.5) - 0.5, linetype = "dashed" ) +


geom_vline( xintercept = which(test_levels == 1), linetype = "dashed" ) +


geom_hline( yintercept = which(retest_levels == 23) - 0.2, linetype = "dashed" ) +


geom_hline( yintercept = which(retest_levels == 23) + 0.5, linetype = "dashed" ) +

scale_size_continuous( name = "Number of strains", range = c(2, 10), breaks = scales::pretty_breaks() ) + scale_x_discrete(drop = FALSE) + scale_y_discrete(drop = FALSE) + scale_colour_manual( name = "Test → Re-test", values = c( "No change" = "#8EC476", "S to I" = "#E5C07B", "S to R" = "#D66D27", "I to S" = "#E5C07B", "I to R" = "#D66D27", "R to S" = "#D66D27", "R to I" = "#E5C07B" ) ) + theme_classic() + labs( x = "Test MIC (mg/L)", y = "Re-test disk diffusion (mm)" )

discordant_mic_disk_3

```

# Concordant phenotype - disk test vs MIC retest
```{r, concordant_disk_mic_4}
cipro_antibiogram_MIC_concordant <- cipro_antibiogram_DD %>%
  filter(Laboratory.Typing.Method == "Disk diffusion") %>%
  filter(ConcordantRetesting_Method == "E-test (EUCAST 2026)") %>%
  select(
    strain,
    Measurement,
    ConcordantRetesting_Measurement
  ) %>%
  filter(!is.na(ConcordantRetesting_Measurement)) %>%
  rename(
    Test = Measurement,
    `Re-test` = ConcordantRetesting_Measurement
  ) %>%
  mutate(
    Test = as.character(Test),
    `Re-test` = as.character(`Re-test`)
  )


# Helper: build numerically sorted ordered level vector
build_sign_aware_levels <- function(labels) {
  labels <- unique(labels)

  order_key <- sapply(
    labels,
    function(x) {
      value <- as.numeric(gsub("[^0-9.]", "", x))

      if (grepl("^<", x)) {
        value - 0.001
      } else if (grepl("^>", x)) {
        value + 0.001
      } else {
        value
      }
    }
  )

  labels[order(order_key)]
}


# Separate level vectors
# Test = disk diffusion
# Re-test = MIC

disk_levels <- build_sign_aware_levels(
  cipro_antibiogram_MIC_concordant$Test
)

mic_levels <- build_sign_aware_levels(
  cipro_antibiogram_MIC_concordant$`Re-test`
)


cipro_antibiogram_MIC_concordant <- 
  cipro_antibiogram_MIC_concordant %>%
  mutate(
    Test = factor(
      Test,
      levels = disk_levels,
      ordered = TRUE
    ),
    `Re-test` = factor(
      `Re-test`,
      levels = mic_levels,
      ordered = TRUE
    )
  )


# Aggregate counts
cipro_antibiogram_MIC_concordant_wide <- 
  cipro_antibiogram_MIC_concordant %>%
  count(
    Test,
    `Re-test`,
    name = "n_strains"
  )


# Classify Test and Re-test
#
# Disk diffusion:
# R = <=21 mm
# I = 22-24 mm
# S = >=25 mm
#
# MIC:
# S = <0.5 mg/L
# I = 0.5 to <1 mg/L
# R = >=1 mg/L

cipro_antibiogram_MIC_concordant_wide <- 
  cipro_antibiogram_MIC_concordant_wide %>%
  mutate(

    # Disk diffusion Test classification
    Test_category = case_when(
      as.numeric(as.character(Test)) <= 21 ~ "R",
      as.numeric(as.character(Test)) >= 22 &
        as.numeric(as.character(Test)) <= 24 ~ "I",
      as.numeric(as.character(Test)) >= 25 ~ "S"
    ),

    # MIC Re-test classification
    Retest_numeric = as.numeric(
      gsub("[^0-9.]", "", as.character(`Re-test`))
    ),

    Retest_category = case_when(
      Retest_numeric < 0.5 ~ "S",
      Retest_numeric >= 0.5 &
        Retest_numeric < 1 ~ "I",
      Retest_numeric >= 1 ~ "R"
    ),

    # Label unchanged results as "No change"
    Transition = ifelse(
      Test_category == Retest_category,
      "No change",
      paste(
        Test_category,
        "to",
        Retest_category
      )
    )
  )
# Create size mapping specifically for the concordant disk-MIC dataset
size_values_disk_mic_concordant <- sort(
  unique(cipro_antibiogram_MIC_concordant_wide$n_strains)
)

size_mapping_disk_mic_concordant <- setNames(
  seq(
    2,
    10,
    length.out = length(size_values_disk_mic_concordant)
  ),
  size_values_disk_mic_concordant
)


# Concordant Disk diffusion vs MIC plot
concordant_disk_mic_4 <-
  ggplot(
    cipro_antibiogram_MIC_concordant_wide,
    aes(
      x = Test,
      y = `Re-test`,
      size = factor(n_strains),
      colour = Transition
    )
  ) +

  geom_point(
    alpha = 0.7
  ) +

  # -------------------------
  # Disk diffusion boundaries
  # -------------------------

  # R/I boundary: 21 | 22 mm
  geom_vline(
    xintercept = which(disk_levels == "21") + 0.70,
    linetype = "dashed"
  ) +

  # I/S boundary: 24 | 25 mm
  geom_vline(
    xintercept = which(disk_levels == "28") - 0.75,
    linetype = "dashed"
  ) +

  # -------------------------
  # MIC boundaries
  # -------------------------

  # S/I boundary: 1 mg/L
  geom_hline(
    yintercept = which(mic_levels == "1"),
    linetype = "dashed"
  ) +

  # Additional MIC boundary: 0.38 mg/L
  geom_hline(
    yintercept = which(mic_levels == "0.38"),
    linetype = "dashed"
  ) +

  scale_x_discrete(
    drop = FALSE
  ) +

  scale_y_discrete(
    drop = FALSE
  ) +

  # One unique circle size for each actual strain count
  scale_size_manual(
    name = "Number of strains",
    values = size_mapping_disk_mic_concordant,
    drop = FALSE
  ) +

  scale_colour_manual(
    name = "Test → Re-test",
    values = c(
      "No change" = "#8EC476",

      "R to I" = "#E5C07B",
      "R to S" = "#D66D27",

      "I to R" = "#E5C07B",
      "I to S" = "#E5C07B",

      "S to I" = "#E5C07B",
      "S to R" = "#D66D27"
    )
  ) +

  theme_classic() +

  labs(
    x = "Test disk diffusion (mm)",
    y = "Re-test MIC (mg/L)"
  )


concordant_disk_mic_4
```

# Concordant phenotype - disk test vs disk retest


```{r, concordant_disk_disk_5}

# Create concordant DD dataset
cipro_antibiogram_DD_concordant <- cipro_antibiogram_DD %>%
  filter(Laboratory.Typing.Method == "Disk diffusion") %>%
  filter(is.na(ConcordantRetesting_Method)) %>%
  select(
    strain,
    Measurement,
    ConcordantRetesting_Measurement
  ) %>%
  filter(!is.na(ConcordantRetesting_Measurement)) %>%
  rename(
    Test = Measurement,
    `Re-test` = ConcordantRetesting_Measurement
  ) %>%
  mutate(
    Test = as.numeric(Test),
    `Re-test` = as.numeric(`Re-test`)
  )


# Build ONE shared level vector for both axes
# Force 21, 22, 24, and 25 into the levels
DD_levels <- sort(unique(c(
  cipro_antibiogram_DD_concordant$Test,
  cipro_antibiogram_DD_concordant$`Re-test`,
  21, 22, 24, 25
)))


# Apply the SAME level vector to both Test and Re-test
cipro_antibiogram_DD_concordant <- 
  cipro_antibiogram_DD_concordant %>%
  mutate(
    Test = factor(
      Test,
      levels = DD_levels,
      ordered = TRUE
    ),
    `Re-test` = factor(
      `Re-test`,
      levels = DD_levels,
      ordered = TRUE
    )
  )


# Aggregate counts
cipro_antibiogram_DD_concordant_wide <- 
  cipro_antibiogram_DD_concordant %>%
  count(
    Test,
    `Re-test`,
    name = "n_strains"
  )


# Classify DD measurements:
# R = <=21 mm
# I = 22-24 mm
# S = >=25 mm

classify_DD <- function(label) {

  numeric_value <- as.numeric(
    gsub("[^0-9.]", "", label)
  )

  case_when(
    numeric_value <= 21 ~ "R",
    numeric_value >= 22 & numeric_value <= 24 ~ "I",
    numeric_value >= 25 ~ "S"
  )
}


# Add Test/Re-test classification and transition
cipro_antibiogram_DD_concordant_wide <- 
  cipro_antibiogram_DD_concordant_wide %>%
  mutate(

    Test_category = classify_DD(
      as.character(Test)
    ),

    Retest_category = classify_DD(
      as.character(`Re-test`)
    ),

    Transition = ifelse(
      Test_category == Retest_category,
      "No change",
      paste(
        Test_category,
        "to",
        Retest_category
      )
    )
  )


# Scatter plot: Test vs Re-test
# Sized by strain count and coloured by Test -> Re-test transition

# Create size mapping specifically for the concordant dataset
size_values_concordant <- sort(
  unique(cipro_antibiogram_DD_concordant_wide$n_strains)
)

size_mapping_concordant <- setNames(
  seq(
    2,
    10,
    length.out = length(size_values_concordant)
  ),
  size_values_concordant
)


# Concordant Test vs Re-test scatter plot
concordant_disk_disk_5 <-
  ggplot(
    cipro_antibiogram_DD_concordant_wide,
    aes(
      x = Test,
      y = `Re-test`,
      size = factor(n_strains),
      colour = Transition
    )
  ) +

  geom_point(
    alpha = 0.7
  ) +

  # R/I boundary: 21 | 22 mm
  geom_vline(
    xintercept = which(DD_levels == 21) + 0.5,
    linetype = "dashed"
  ) +

  geom_hline(
    yintercept = which(DD_levels == 21) + 0.5,
    linetype = "dashed"
  ) +

  # I/S boundary: 24 | 25 mm
  geom_vline(
    xintercept = which(DD_levels == 24) + 0.5,
    linetype = "dashed"
  ) +

  geom_hline(
    yintercept = which(DD_levels == 24) + 0.5,
    linetype = "dashed"
  ) +

  scale_x_discrete(
    drop = FALSE
  ) +

  scale_y_discrete(
    drop = FALSE
  ) +

  # One unique circle size for each actual strain count
  scale_size_manual(
    name = "Number of strains",
    values = size_mapping_concordant,
    drop = FALSE
  ) +

  scale_colour_manual(
    name = "Test → Re-test",
    values = c(
      "No change" = "#8EC476",

      "R to I" = "#E5C07B",
      "I to R" = "#E5C07B",

      "I to S" = "#E5C07B",
      "S to I" = "#E5C07B",

      "R to S" = "#D66D27",
      "S to R" = "#D66D27"
    )
  ) +

  theme_classic() +

  labs(
    x = "Test disk diffusion (mm)",
    y = "Re-test disk diffusion (mm)"
  )


concordant_disk_disk_5
```


# Figure S3 - MIC and disk distribution
``` {r figs3}

((discordant_mic_mic_1 + ggtitle(label="a) Re-testing discordant phenotypes", subtitle = "MIC test vs. MIC re-test")) | discordant_disk_disk_2 +ggtitle(label="", subtitle = "Disk test vs. Disk re-test")) /
  (concordant_disk_disk_5 + ggtitle(label="b) Re-testing concordant phenotypes", subtitle = "Disk test vs. Disk re-test")|concordant_disk_mic_4 +ggtitle(label= "", subtitle = "Disk test vs. MIC re-test"))



ggsave(width=19, height=10, file="figs/FigS3_retest_distribution.pdf")
ggsave(width=19, height=10, file="figs/FigS3_retest_distribution.png")

```