---
title: "Validation"
author: "Kara Tsang, Kat Holt"
date: "Last compiled on `r format(Sys.time(), '%d/%m/%y')`"
output: 
  html_document: 
    df_print: paged
    fig_width: 9
    fig_height: 8
    mode: selfcontained
    highlight: pygments
    theme: 
      bootswatch: flatly
      code_font:
        google: "JetBrains Mono"
editor_options: 
  chunk_output_type: inline
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 1600px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = 'png', echo = TRUE, message = FALSE, warning = FALSE,
                      cache = FALSE, fig.align = 'center')
knitr::opts_knit$set(root.dir = getwd())

# Data wrangling
library(tidyverse)
library(dplyr)
library(janitor)
library(png)
library(patchwork)
```

# load data
``` {r load_data}
MBIRA_Kpn <- read_delim("MBIRA_ValidationOutput/Prediction_Pheno_summary_Kpn.csv")
MBIRA_Kvv <- read_delim("MBIRA_ValidationOutput/Prediction_Pheno_summary_Kvv.csv")
HowdenGorrie <- read_delim("HowdenGorrie_ValidationOutput/Prediction_Pheno_summary.csv")
AGAR <- read_delim("AGAR_ValidationOutput/Prediction_Pheno_summary.csv")
Milan <- read_delim("MilanSanRafaelle_ValidationOutput/Prediction_Pheno_summary.csv")
HailuEthiopia <- read_delim("HailuEthiopia_ValidationOutput/Prediction_Pheno_summary.csv")
FeaseyMalawi <- read_delim("FeaseyMalawi_ValidationOutput/Prediction_Pheno_summary.csv")
KayamaJapan <- read_delim("KayamaJapan_ValidationOutput/Prediction_Pheno_summary.csv")
EUSCAPE_EURECA_Kpn <- read_delim("EURECA_EUSCAPE_ValidationOutput/Prediction_Pheno_summary_Kpn.csv")
EUSCAPE_EURECA_Kvv <- read_delim("EURECA_EUSCAPE_ValidationOutput/Prediction_Pheno_summary_Kvv.csv")
EgliSwitzerland <- read_delim("EgliSwitzerland_2025_ValidationOutput/Prediction_Pheno_summary.csv")
HRYC_Spain <- read_delim("HRYC_Spain_ValidationOutput/Prediction_Pheno_summary.csv")
IndiaGHRU <- read_delim("IndiaGHRU_2025_ValidationOutput/Prediction_Pheno_summary.csv")
NeumannGermany <- read_delim("NeumannGermany_ValidationOutput/Prediction_Pheno_summary.csv")
Whitelaw_CHAMPS <- read_delim("Whitelaw_Stellenbosch_ValidationOutput/Prediction_Pheno_summary_DD.csv")
Oxfordshire_Kpn <- read_delim("Oxfordshire_ValidationOutput/Prediction_Pheno_summary_Kpn.csv")
Oxfordshire_Kvv <- read_delim("Oxfordshire_ValidationOutput/Prediction_Pheno_summary_Kvv.csv")
IZSAM_Italy <- read_delim("IZSAM_Italy_ValidationOutput/Prediction_Pheno_summary.csv")
LaPaz_Lazaro <- read_delim("LaPaz_Lazaro_ValidationOutput/Prediction_Pheno_summary.csv")
HarrisMERINO <- read_delim("HarrisMERINO_ValidationOutput/Prediction_Pheno_summary.csv")
PachecoBrazil <- read_delim("PachecoBrazil_ValidationOutput/Prediction_Pheno_summary.csv")
SmaouiTunisia <- read_delim("SmaouiTunisia_ValidationOutput/Prediction_Pheno_summary.csv")
```

## set colours
``` {r set_colours}
wt_colours <- c(`non-wt (I/R)`="IndianRed", `wt (S)`= "LightBlue")
res_colours <- c("I"="grey", "R"="IndianRed", "S"="LightBlue", "NWT"="grey")
res_colours2 <- c("I"="black", "R"="IndianRed", "S"="LightBlue", "NWT"="black")
res_colours3 <- c("NWT I"="#78638a", "NWT R"="IndianRed", "WT S"="LightBlue")
```

``` {r}
# position dodge for coefficient plots
pd <- position_dodge(width=0.8)
```

```{r metrics}
## functions to calculate PPV and categorical agreement from models

prop <- function(x, n) {
  if (n == 0) return(NA_real_)
  x / n
}

prop_ci <- function(p, n, z = 1.96) {
  if (n == 0 || is.na(p)) {
    return(c(lower = NA_real_, upper = NA_real_))
  }

  se <- sqrt((p * (1 - p)) / n)

  c(
    lower = p - z * se,
    upper = p + z * se
  )
}

get_cells <- function(xtabs) {

  stopifnot(all(dim(xtabs) == c(2,2)))

  list(
    TN = xtabs[1,1],
    FN = xtabs[1,2],
    FP = xtabs[2,1],
    TP = xtabs[2,2]
  )
}

# ---- minor error from 3x3 S/I/R table ----

minor_error_I <- function(xtabs3) {

  xt3 <- as.data.frame(xtabs3)

  # detect row-name column from imported tables
  idx_col <- which(
    names(xt3) %in% c("...1", "..1") |
      grepl("^\\.\\.", names(xt3))
  )

  # if row labels stored in first column, move to rownames
  if (length(idx_col) > 0) {

    row_ids <- as.character(xt3[[idx_col[1]]])

    xt3 <- xt3[, -idx_col, drop = FALSE]

    rownames(xt3) <- row_ids
  }

  # ensure required rows/columns exist
  required_rows <- c("pred_S", "pred_I", "pred_R")
  required_cols <- c("S", "I", "R")

  missing_rows <- setdiff(required_rows, rownames(xt3))
  missing_cols <- setdiff(required_cols, colnames(xt3))

  if (length(missing_rows) > 0) {
    stop(
      paste("Missing rows in xtabs3:",
            paste(missing_rows, collapse = ", "))
    )
  }

  if (length(missing_cols) > 0) {
    stop(
      paste("Missing columns in xtabs3:",
            paste(missing_cols, collapse = ", "))
    )
  }

  # reorder safely
  xt3 <- xt3[required_rows, required_cols, drop = FALSE]

  xt3 <- as.matrix(xt3)
  storage.mode(xt3) <- "numeric"

  N <- sum(xt3, na.rm = TRUE)

  me_count <-
    xt3["pred_I", "S"] +
    xt3["pred_I", "R"] +
    xt3["pred_S", "I"] +
    xt3["pred_R", "I"]

  p <- ifelse(N == 0, NA_real_, me_count / N)

  ci <- prop_ci(p, N)

  list(
    me_I_total = p,
    me_I_total_lb = unname(ci["lower"]),
    me_I_total_ub = unname(ci["upper"])
  )
}

# ---- main metrics function ----

metrics <- function(xtabs, xtabs3 = NULL) {

  x <- get_cells(xtabs)

  n <- sum(xtabs)

  # core metrics
  cat  <- prop(x$TN + x$TP, n)
  sens <- prop(x$TP, x$TP + x$FN)
  spec <- prop(x$TN, x$TN + x$FP)
  ppv  <- prop(x$TP, x$TP + x$FP)
  npv  <- prop(x$TN, x$TN + x$FN)
  me   <- prop(x$FP, x$TN + x$FP)
  vme  <- prop(x$FN, x$TP + x$FN)

  perc_R <- prop(x$TP + x$FN, n)

  # confidence intervals
  cat_ci  <- prop_ci(cat, n)
  sens_ci <- prop_ci(sens, x$TP + x$FN)
  spec_ci <- prop_ci(spec, x$TN + x$FP)
  ppv_ci  <- prop_ci(ppv, x$TP + x$FP)
  npv_ci  <- prop_ci(npv, x$TN + x$FN)
  me_ci   <- prop_ci(me, x$TN + x$FP)
  vme_ci  <- prop_ci(vme, x$TP + x$FN)

  ci_low  <- function(ci) unname(ci["lower"])
  ci_high <- function(ci) unname(ci["upper"])

  # ---- minor error ----

  if (!is.null(xtabs3)) {

    me_I <- minor_error_I(xtabs3)

    me_I_total <- me_I$me_I_total

    me_I_total_ci <- c(
      lower = me_I$me_I_total_lb,
      upper = me_I$me_I_total_ub
    )

  } else {

    me_I_total <- NA_real_

    me_I_total_ci <- c(
      lower = NA_real_,
      upper = NA_real_
    )
  }

  summary <- c(

    cat = cat,
    sens = sens,
    spec = spec,
    me = me,
    vme = vme,

    cat_lb = ci_low(cat_ci),
    cat_ub = ci_high(cat_ci),

    sens_lb = ci_low(sens_ci),
    sens_ub = ci_high(sens_ci),

    spec_lb = ci_low(spec_ci),
    spec_ub = ci_high(spec_ci),

    me_lb = ci_low(me_ci),
    me_ub = ci_high(me_ci),

    vme_lb = ci_low(vme_ci),
    vme_ub = ci_high(vme_ci),

    perc_R = perc_R,

    me_I_total = me_I_total,
    me_I_total_lb = ci_low(me_I_total_ci),
    me_I_total_ub = ci_high(me_I_total_ci),

    n = n
  )

  list(

    cat = cat,
    cat_lb = ci_low(cat_ci),
    cat_ub = ci_high(cat_ci),

    ppv = ppv,
    ppv_lb = ci_low(ppv_ci),
    ppv_ub = ci_high(ppv_ci),

    npv = npv,
    npv_lb = ci_low(npv_ci),
    npv_ub = ci_high(npv_ci),

    sens = sens,
    sens_lb = ci_low(sens_ci),
    sens_ub = ci_high(sens_ci),

    spec = spec,
    spec_lb = ci_low(spec_ci),
    spec_ub = ci_high(spec_ci),

    me = me,
    me_lb = ci_low(me_ci),
    me_ub = ci_high(me_ci),

    vme = vme,
    vme_lb = ci_low(vme_ci),
    vme_ub = ci_high(vme_ci),

    perc_R = perc_R,

    me_I_total = me_I_total,
    me_I_total_lb = ci_low(me_I_total_ci),
    me_I_total_ub = ci_high(me_I_total_ci),

    n = n
  )
}
```

```{r process_prediction}
process_prediction <- function(dat) {
  
  dat_SI <- dat %>% 
    rename(Group = ...1) %>% 
    filter(Group %in% c("pred_S", "pred_I")) %>%
    rowwise() %>%
    mutate(
      S_I = sum(S, I),
      I_R = sum(I, R)
    ) %>%
    select(-I) %>%
    adorn_totals("row") %>%
    mutate(Group = gsub("Total", "pred_S_I", Group))
  
  dat_R <- dat %>% 
    rename(Group = ...1) %>% 
    filter(Group %in% c("pred_R", "pred_I")) %>%
    rowwise() %>%
    mutate(
      S_I = sum(S, I),
      I_R = sum(I, R)
    ) %>%
    select(-I) %>%
    adorn_totals("row") %>%
    mutate(Group = gsub("Total", "pred_I_R", Group))
  
  dat_SIR <- rbind(dat_SI, dat_R) %>% distinct()
  
dat_3 <- dat[1:3, 1:4]

# remove first column and convert to matrix
dat_3_matrix <- as.matrix(dat_3[, 2:4])

storage.mode(dat_3_matrix) <- "numeric"

# explicitly assign row/column names
rownames(dat_3_matrix) <- c("pred_S", "pred_I", "pred_R")
colnames(dat_3_matrix) <- c("S", "I", "R")

dat_3 <- dat_3_matrix
  
  minor_error_I(dat_3)
  
  subset(dat_SIR, Group == "pred_S_I")$S_I
  table(dat_SIR$S_I)
  
  R_prediction <- matrix(
    c(
      subset(dat_SIR, Group == "pred_S_I")$S_I,
      subset(dat_SIR, Group == "pred_S_I")$R,
      subset(dat_SIR, Group == "pred_R")$S_I,
      subset(dat_SIR, Group == "pred_R")$R
    ),
    ncol = 2,
    byrow = TRUE
  )
  
  colnames(R_prediction) <- c("S_I", "R")
  rownames(R_prediction) <- c("pred_S_I", "pred_R")
  
  R_prediction_summary <- metrics(R_prediction)
  R_prediction_summary <- metrics(R_prediction, dat_3)
  
  NWT_prediction <- matrix(
    c(
      subset(dat_SIR, Group == "pred_S")$S,
      subset(dat_SIR, Group == "pred_S")$I_R,
      subset(dat_SIR, Group == "pred_I_R")$S,
      subset(dat_SIR, Group == "pred_I_R")$I_R
    ),
    ncol = 2,
    byrow = TRUE
  )
  
  colnames(NWT_prediction) <- c("S", "I_R")
  rownames(NWT_prediction) <- c("pred_S", "pred_I_R")
  
  NWT_prediction_summary <- metrics(NWT_prediction, dat_3)
  
  list(
    SIR = dat_SIR,
    dat_3 = dat_3,
    R_prediction = R_prediction,
    R_prediction_summary = R_prediction_summary,
    NWT_prediction = NWT_prediction,
    NWT_prediction_summary = NWT_prediction_summary
  )
}

```

```{r process_results}

MBIRA_Kpn_results <- process_prediction(MBIRA_Kpn)
MBIRA_Kvv_results <- process_prediction(MBIRA_Kvv)
HowdenGorrie_results <- process_prediction(HowdenGorrie)
AGAR_results <- process_prediction(AGAR)
Milan_results <- process_prediction(Milan)
HailuEthiopia_results <- process_prediction(HailuEthiopia)
KayamaJapan_results <- process_prediction(KayamaJapan)
FeaseyMalawi_results <- process_prediction(FeaseyMalawi)
EUSCAPE_EURECA_Kpn_results <- process_prediction(EUSCAPE_EURECA_Kpn)
EUSCAPE_EURECA_Kvv_results <- process_prediction(EUSCAPE_EURECA_Kvv)
EgliSwitzerland_results <- process_prediction(EgliSwitzerland)
HRYC_Spain_results <- process_prediction(HRYC_Spain)
IndiaGHRU_results <- process_prediction(IndiaGHRU)
NeumannGermany_results <- process_prediction(NeumannGermany)
Whitelaw_CHAMPS_results <- process_prediction(Whitelaw_CHAMPS)
Oxfordshire_Kpn_results <- process_prediction(Oxfordshire_Kpn)
Oxfordshire_Kvv_results <- process_prediction(Oxfordshire_Kvv)
IZSAM_Italy_results <- process_prediction(IZSAM_Italy)
LaPaz_Lazaro_results <- process_prediction(LaPaz_Lazaro)
HarrisMERINO_results <- process_prediction(HarrisMERINO)
PachecoBrazil_results <- process_prediction(PachecoBrazil)
SmaouiTunisia_results <- process_prediction(SmaouiTunisia)
```


```{r pooledR}
pooled_R <- 
  bind_rows(
    MBIRA_Kpn_results$SIR,
    MBIRA_Kvv_results$SIR, 
    HowdenGorrie_results$SIR,
    AGAR_results$SIR,
    Milan_results$SIR,
    HailuEthiopia_results$SIR,
    KayamaJapan_results$SIR,
    FeaseyMalawi_results$SIR, 
    EUSCAPE_EURECA_Kpn_results$SIR,
    EUSCAPE_EURECA_Kvv_results$SIR,
    EgliSwitzerland_results$SIR,
    HRYC_Spain_results$SIR,
    IndiaGHRU_results$SIR,
    NeumannGermany_results$SIR,
    Whitelaw_CHAMPS_results$SIR,
    Oxfordshire_Kpn_results$SIR,
    Oxfordshire_Kvv_results$SIR,
    IZSAM_Italy_results$SIR, 
    LaPaz_Lazaro_results$SIR,
    HarrisMERINO_results$SIR,
    PachecoBrazil_results$SIR,
    SmaouiTunisia_results$SIR
  ) %>%
  group_by(Group) %>%
  summarise_all(~sum(., na.rm = TRUE))

pooled_R_prediction <- matrix(
  c(
    subset(pooled_R, Group == "pred_S_I")$S_I, 
    subset(pooled_R, Group == "pred_S_I")$R,
    subset(pooled_R, Group == "pred_R")$S_I,
    subset(pooled_R, Group == "pred_R")$R
  ),
  ncol = 2,
  byrow = TRUE
)

colnames(pooled_R_prediction) <- c("S_I", "R")
rownames(pooled_R_prediction) <- c("pred_S_I", "pred_R")

pooled_3 <- Reduce(
  "+",
  list(
    MBIRA_Kpn_results$dat_3,
    MBIRA_Kvv_results$dat_3,
    HowdenGorrie_results$dat_3,
    AGAR_results$dat_3,
    Milan_results$dat_3,
    HailuEthiopia_results$dat_3,
    KayamaJapan_results$dat_3,
    FeaseyMalawi_results$dat_3,
    EUSCAPE_EURECA_Kpn_results$dat_3,
    EUSCAPE_EURECA_Kvv_results$dat_3,
    EgliSwitzerland_results$dat_3,
    HRYC_Spain_results$dat_3,
    IndiaGHRU_results$dat_3,
    NeumannGermany_results$dat_3,
    Whitelaw_CHAMPS_results$dat_3,
    Oxfordshire_Kpn_results$dat_3,
    Oxfordshire_Kvv_results$dat_3,
    IZSAM_Italy_results$dat_3,
    LaPaz_Lazaro_results$dat_3,
    HarrisMERINO_results$dat_3,
    PachecoBrazil_results$dat_3,
    SmaouiTunisia_results$dat_3
  )
)

pooled_R_summary <- metrics(pooled_R_prediction, pooled_3)
```

# summary of pooled genotypes vs. SIR
```{r genotypes_SIR}
EUSCAPE_EURECA_Kpn_genotype <- EUSCAPE_EURECA_Kpn %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

EUSCAPE_EURECA_Kvv_genotype <- EUSCAPE_EURECA_Kvv %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R")

KayamaJapan_genotype <- KayamaJapan %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R")

IndiaGHRU_genotype <- IndiaGHRU %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

AGAR_genotype <- AGAR %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

Oxfordshire_Kpn_genotype <-Oxfordshire_Kpn %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

Oxfordshire_Kvv_genotype <-Oxfordshire_Kvv %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

MBIRA_Kpn_genotype <- MBIRA_Kpn %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

HowdenGorrie_genotype <- HowdenGorrie %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

Milan_genotype <- Milan %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

HRYC_Spain_genotype <- HRYC_Spain %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

LaPaz_Lazaro_genotype <- LaPaz_Lazaro %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R")

MBIRA_Kvv_genotype <- MBIRA_Kvv %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

HarrisMERINO_genotype <- HarrisMERINO %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

PachecoBrazil_genotype <- PachecoBrazil %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 
IZSAM_Italy_genotype <- IZSAM_Italy %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

HailuEthiopia_genotype <- HailuEthiopia %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 
FeaseyMalawi_genotype <- FeaseyMalawi %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 


EgliSwitzerland_genotype <- EgliSwitzerland %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 


NeumannGermany_genotype <- NeumannGermany %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 
Whitelaw_CHAMPS_genotype <- Whitelaw_CHAMPS %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 

SmaouiTunisia_genotype <- SmaouiTunisia %>% 
  rename(Group=...1) %>% 
  filter(Group!="pred_S" & Group!="pred_I"& Group!="pred_R") 


pooled_genotype <- 
bind_rows(MBIRA_Kpn_genotype,
          MBIRA_Kvv_genotype, 
          HowdenGorrie_genotype,
          AGAR_genotype,
          Milan_genotype,
          HailuEthiopia_genotype,
          KayamaJapan_genotype,
          FeaseyMalawi_genotype, 
          EUSCAPE_EURECA_Kpn_genotype,
          EUSCAPE_EURECA_Kvv_genotype,
          EgliSwitzerland_genotype,
          HRYC_Spain_genotype,
          IndiaGHRU_genotype,
          NeumannGermany_genotype,
          Whitelaw_CHAMPS_genotype,
          Oxfordshire_Kpn_genotype,
          Oxfordshire_Kvv_genotype,
          IZSAM_Italy_genotype, 
          LaPaz_Lazaro_genotype,
          HarrisMERINO_genotype,
          PachecoBrazil_genotype,
          SmaouiTunisia_genotype) %>%
    group_by(Group) %>%
    summarise_all(funs(sum(., na.rm = TRUE)))

```

```{r}
pooled_NWT <- 
  bind_rows(
    MBIRA_Kpn_results$SIR,
    MBIRA_Kvv_results$SIR, 
    HowdenGorrie_results$SIR,
    AGAR_results$SIR,
    Milan_results$SIR,
    HailuEthiopia_results$SIR,
    KayamaJapan_results$SIR,
    FeaseyMalawi_results$SIR, 
    EUSCAPE_EURECA_Kpn_results$SIR,
    EUSCAPE_EURECA_Kvv_results$SIR,
    EgliSwitzerland_results$SIR,
    HRYC_Spain_results$SIR,
    IndiaGHRU_results$SIR,
    NeumannGermany_results$SIR,
    Whitelaw_CHAMPS_results$SIR,
    Oxfordshire_Kpn_results$SIR,
    Oxfordshire_Kvv_results$SIR,
    IZSAM_Italy_results$SIR, 
    LaPaz_Lazaro_results$SIR,
    HarrisMERINO_results$SIR,
    PachecoBrazil_results$SIR,
    SmaouiTunisia_results$SIR
  ) %>%
  group_by(Group) %>%
  summarise_all(~sum(., na.rm = TRUE))

pooled_NWT_prediction <- matrix(
  c(
    subset(pooled_NWT, Group == "pred_S")$S, 
    subset(pooled_NWT, Group == "pred_S")$I_R,
    subset(pooled_NWT, Group == "pred_I_R")$S,
    subset(pooled_NWT, Group == "pred_I_R")$I_R
  ),
  ncol = 2,
  byrow = TRUE
)

colnames(pooled_NWT_prediction) <- c("S", "I_R")
rownames(pooled_NWT_prediction) <- c("pred_S", "pred_I_R")

pooled_3 <-  Reduce(
  "+",
  list(
  MBIRA_Kpn_results$dat_3,
  MBIRA_Kvv_results$dat_3,
  HowdenGorrie_results$dat_3,
  AGAR_results$dat_3,
  Milan_results$dat_3,
  HailuEthiopia_results$dat_3,
  KayamaJapan_results$dat_3,
  FeaseyMalawi_results$dat_3,
  EUSCAPE_EURECA_Kpn_results$dat_3,
  EUSCAPE_EURECA_Kvv_results$dat_3,
  EgliSwitzerland_results$dat_3,
  HRYC_Spain_results$dat_3,
  IndiaGHRU_results$dat_3,
  NeumannGermany_results$dat_3,
  Whitelaw_CHAMPS_results$dat_3,
  Oxfordshire_Kpn_results$dat_3,
  Oxfordshire_Kvv_results$dat_3,
  IZSAM_Italy_results$dat_3,
  LaPaz_Lazaro_results$dat_3,
  HarrisMERINO_results$dat_3,
  PachecoBrazil_results$dat_3,
  SmaouiTunisia_results$dat_3
))

pooled_NWT_summary <- metrics(pooled_NWT_prediction, pooled_3)
```

# TOTAL SIR counts per dataset
```{r}
SIR_counts_table <- bind_rows(lst(MBIRA_Kpn_results$SIR,
    MBIRA_Kvv_results$SIR, 
    HowdenGorrie_results$SIR,
    AGAR_results$SIR,
    Milan_results$SIR,
    HailuEthiopia_results$SIR,
    KayamaJapan_results$SIR,
    FeaseyMalawi_results$SIR, 
    EUSCAPE_EURECA_Kpn_results$SIR,
    EUSCAPE_EURECA_Kvv_results$SIR,
    EgliSwitzerland_results$SIR,
    HRYC_Spain_results$SIR,
    IndiaGHRU_results$SIR,
    NeumannGermany_results$SIR,
    Whitelaw_CHAMPS_results$SIR,
    Oxfordshire_Kpn_results$SIR,
    Oxfordshire_Kvv_results$SIR,
    IZSAM_Italy_results$SIR, 
    LaPaz_Lazaro_results$SIR,
    HarrisMERINO_results$SIR,
    PachecoBrazil_results$SIR,
    SmaouiTunisia_results$SIR), .id = 'Dataset') %>%
  filter(Group!="pred_S_I" & Group!="pred_I_R")

SIR_counts_table$Dataset <- gsub("\\_results\\$SIR",'',SIR_counts_table$Dataset)

SIR_counts_table<-SIR_counts_table %>% select(-Group, -S_I)

S_counts_table <-
SIR_counts_table %>% 
  group_by(Dataset) %>% 
  summarise(S_total = sum(S))

R_counts_table <-
SIR_counts_table %>% 
  group_by(Dataset) %>% 
  summarise(R_total = sum(R))

IR_counts_table <-
  SIR_counts_table %>%
  group_by(Dataset) %>% 
  summarise(IR_total = sum(I_R))

SR_table<- full_join(S_counts_table, R_counts_table) 

SIR_table<- full_join(SR_table, IR_counts_table)  %>% 
  mutate(I_total=IR_total-R_total) %>% 
  mutate(N_total=S_total+IR_total)

SIR_table$Dataset[SIR_table$Dataset == 'AGAR'] <- 'AGAR GnSOP'
SIR_table$Dataset[SIR_table$Dataset == 'EUSCAPE_EURECA_Kpn'] <- 'EUSCAPE & EURECA'
SIR_table$Dataset[SIR_table$Dataset == 'EUSCAPE_EURECA_Kvv'] <- 'EUSCAPE & EURECA*'
SIR_table$Dataset[SIR_table$Dataset == 'EgliSwitzerland'] <- 'University of Zurich'
SIR_table$Dataset[SIR_table$Dataset == 'FeaseyMalawi'] <- 'Queen Elizabeth Central Hospital'
SIR_table$Dataset[SIR_table$Dataset == 'HRYC_Spain'] <- 'HURYC & CIBERINFEC'
SIR_table$Dataset[SIR_table$Dataset == 'HailuEthiopia'] <- 'Addis Ababa University*'
SIR_table$Dataset[SIR_table$Dataset == 'HowdenGorrie'] <- 'Controlling Superbugs study & Victorian CPE program'
SIR_table$Dataset[SIR_table$Dataset == 'IZSAM_Italy'] <- 'IZSAM'
SIR_table$Dataset[SIR_table$Dataset == 'IndiaGHRU'] <- 'India GHRU'
SIR_table$Dataset[SIR_table$Dataset == 'KayamaJapan'] <- 'JARBS-GNR'
SIR_table$Dataset[SIR_table$Dataset == 'LaPaz_Lazaro'] <- 'HULP'
SIR_table$Dataset[SIR_table$Dataset == 'LaPaz_Lazaro'] <- 'HULP'
SIR_table$Dataset[SIR_table$Dataset == 'MBIRA_Kpn'] <- 'MBIRA'
SIR_table$Dataset[SIR_table$Dataset == 'MBIRA_Kvv'] <- 'MBIRA*'
SIR_table$Dataset[SIR_table$Dataset == 'Milan'] <- 'IRCCS Ospedale San Raffaele'
SIR_table$Dataset[SIR_table$Dataset == 'NeumannGermany'] <- 'SCHARKI'
SIR_table$Dataset[SIR_table$Dataset == 'Oxfordshire_Kpn'] <- 'Oxfordshire'
SIR_table$Dataset[SIR_table$Dataset == 'Oxfordshire_Kvv'] <- 'Oxfordshire*'
SIR_table$Dataset[SIR_table$Dataset == 'Whitelaw_CHAMPS'] <- 'CHAMPS'
SIR_table$Dataset[SIR_table$Dataset == 'SmaouiTunisia'] <- 'SmaouiTunisia*'

```


```{r}
validation_summary <- rbind(
  EUSCAPE_EURECA_Kpn_results$R_prediction_summary,
  EUSCAPE_EURECA_Kpn_results$NWT_prediction_summary,
  
  KayamaJapan_results$R_prediction_summary,
  KayamaJapan_results$NWT_prediction_summary,
  
  IndiaGHRU_results$R_prediction_summary,
  IndiaGHRU_results$NWT_prediction_summary,
  
  AGAR_results$R_prediction_summary,
  AGAR_results$NWT_prediction_summary,
  
  Oxfordshire_Kpn_results$R_prediction_summary,
  Oxfordshire_Kpn_results$NWT_prediction_summary,
  
  IZSAM_Italy_results$R_prediction_summary,
  IZSAM_Italy_results$NWT_prediction_summary,
  
  MBIRA_Kpn_results$R_prediction_summary,
  MBIRA_Kpn_results$NWT_prediction_summary,
  
  HowdenGorrie_results$R_prediction_summary,
  HowdenGorrie_results$NWT_prediction_summary,
  
  Milan_results$R_prediction_summary,
  Milan_results$NWT_prediction_summary,
  
  HRYC_Spain_results$R_prediction_summary,
  HRYC_Spain_results$NWT_prediction_summary,
  
  LaPaz_Lazaro_results$R_prediction_summary,
  LaPaz_Lazaro_results$NWT_prediction_summary,
  
  MBIRA_Kvv_results$R_prediction_summary,
  MBIRA_Kvv_results$NWT_prediction_summary,
  
  Oxfordshire_Kvv_results$R_prediction_summary,
  Oxfordshire_Kvv_results$NWT_prediction_summary,
  
  NeumannGermany_results$R_prediction_summary,
  NeumannGermany_results$NWT_prediction_summary,
  
  EUSCAPE_EURECA_Kvv_results$R_prediction_summary,
  EUSCAPE_EURECA_Kvv_results$NWT_prediction_summary,
  
  PachecoBrazil_results$R_prediction_summary,
  PachecoBrazil_results$NWT_prediction_summary,
  
  HarrisMERINO_results$R_prediction_summary,
  HarrisMERINO_results$NWT_prediction_summary,
  
  FeaseyMalawi_results$R_prediction_summary,
  FeaseyMalawi_results$NWT_prediction_summary,
  
  SmaouiTunisia_results$R_prediction_summary,
  SmaouiTunisia_results$NWT_prediction_summary,
  
  EgliSwitzerland_results$R_prediction_summary,
  EgliSwitzerland_results$NWT_prediction_summary,
  
  HailuEthiopia_results$R_prediction_summary,
  HailuEthiopia_results$NWT_prediction_summary,
  
  Whitelaw_CHAMPS_results$R_prediction_summary,
  Whitelaw_CHAMPS_results$NWT_prediction_summary,
  
  pooled_R_summary,
  pooled_NWT_summary
)


validation_summary_table <- as_tibble(validation_summary) %>%
  
  mutate(
    Prediction = c(
      "R", "NWT","R", "NWT","R", "NWT","R", "NWT",
      "R", "NWT","R", "NWT","R", "NWT","R", "NWT",
      "R", "NWT","R", "NWT","R", "NWT","R", "NWT",
      "R", "NWT","R", "NWT","R", "NWT","R", "NWT",
      "R", "NWT","R", "NWT","R", "NWT","R", "NWT",
      "R", "NWT","R", "NWT","R", "NWT"
    )
  ) %>%
  
  mutate(
    Dataset = c(
      "EUSCAPE & EURECA", "EUSCAPE & EURECA",
      "JARBS-GNR", "JARBS-GNR",
      "India GHRU", "India GHRU",
      "AGAR GnSOP", "AGAR GnSOP",
      "Oxfordshire", "Oxfordshire",
      "IZSAM", "IZSAM",
      "MBIRA", "MBIRA",
      "Controlling Superbugs study & Victorian CPE program",
      "Controlling Superbugs study & Victorian CPE program",
      "IRCCS Ospedale San Raffaele",
      "IRCCS Ospedale San Raffaele",
      "HURYC & CIBERINFEC",
      "HURYC & CIBERINFEC",
      "HULP", "HULP",
      "MBIRA*", "MBIRA*",
      "Oxfordshire*", "Oxfordshire*",
      "SCHARKI", "SCHARKI",
      "EUSCAPE & EURECA*", "EUSCAPE & EURECA*",
      "PachecoBrazil", "PachecoBrazil",
      "HarrisMERINO", "HarrisMERINO",
      "Queen Elizabeth Central Hospital",
      "Queen Elizabeth Central Hospital",
      "SmaouiTunisia*", "SmaouiTunisia*",
      "University of Zurich", "University of Zurich",
      "Addis Ababa University*", "Addis Ababa University*",
      "CHAMPS", "CHAMPS",
      "Pooled", "Pooled"
    )
  ) %>%
  
  mutate(
    Country = c(
      "Various European countries", "Various European countries",
      "Japan", "Japan",
      "India", "India",
      "Australia", "Australia",
      "United Kingdom", "United Kingdom",
      "Italy", "Italy",
      "Tanzania, Ethiopia, Malawi, Zambia, Ghana, Nigeria, South Africa",
      "Tanzania, Ethiopia, Malawi, Zambia, Ghana, Nigeria, South Africa",
      "Australia", "Australia",
      "Italy", "Italy",
      "Spain, Portugal", "Spain, Portugal",
      "Spain", "Spain",
      "Ethiopia, South Africa, Tanzania",
      "Ethiopia, South Africa, Tanzania",
      "United Kingdom", "United Kingdom",
      "Germany", "Germany",
      "Various European countries", "Various European countries",
      "Brazil", "Brazil",
      "Various countries", "Various countries",
      "Malawi", "Malawi",
      "Tunisia", "Tunisia",
      "Switzerland", "Switzerland",
      "Ethiopia", "Ethiopia",
      "South Africa", "South Africa",
      "", ""
    )
  ) %>%
  
  rename(
    Sensitivity = sens,
    Specificity = spec,
    `ME` = me,
    `VME` = vme,
    `Minor errors` = me_I_total,
    `Minerror_lb` = me_I_total_lb,
    `Minerror_ub` = me_I_total_ub,
    `Categorical agreement` = cat,
    `%R` = perc_R,
    N = n
  ) %>%
  
  mutate(
    Sensitivity = paste0(round(as.numeric(Sensitivity) * 100, 2), "%"),
    Specificity = paste0(round(as.numeric(Specificity) * 100, 2), "%"),
    ME = paste0(round(as.numeric(ME) * 100, 2), "%"),
    VME = paste0(round(as.numeric(VME) * 100, 2), "%"),
    `Minor errors` = paste0(round(as.numeric(`Minor errors`) * 100, 2), "%"),
    `Categorical agreement` = paste0(round(as.numeric(`Categorical agreement`) * 100, 2), "%"),
    `%R` = paste0(round(as.numeric(`%R`) * 100, 2), "%")
  )

#write_csv(validation_summary_table, "../tables/Table3_extValidation_evalMetrics.csv", na="-")

```



# Plotting validation sets (only R_excludeI)

```{r plot}


validation_summary_table_plot <- validation_summary_table %>% 
  filter(Prediction =="R") %>%
  relocate(Dataset, Prediction, `Categorical agreement`, `Sensitivity`, `Specificity`, `ME`, `VME`, `Minor errors`)

validation_summary_table_long <- gather(
  validation_summary_table_plot,
  `Evaluation metric`,
  value,
  `Categorical agreement`:`Minor errors`
)

validation_summary_table_long <- 
  validation_summary_table_long %>%
  mutate(value = as.numeric(gsub("%", "", value))) %>%
  rename(metric_value = value)

validation_summary_table_long_ci <-
  validation_summary_table_long %>% 
  gather(v, ci_value, cat_lb:Minerror_ub) %>% 
  separate(v, c("var", "col")) %>% 
  arrange(Dataset) %>% 
  pivot_wider(
    names_from = col,
    values_from = ci_value
  )

validation_summary_table_long_ci <- 
  validation_summary_table_long_ci %>% 
  filter(
    `Evaluation metric` == "Categorical agreement" & var == "cat" |
    `Evaluation metric` == "ME" & var == "me" |
    `Evaluation metric` == "Minor errors" & var == "Minerror" |
    `Evaluation metric` == "VME" & var == "vme"
  )


validation_summary_table_total <- 
  full_join(validation_summary_table_long_ci, SIR_table) %>% 
  mutate(Dataset = gsub("Pooled", "All datasets combined", Dataset)) %>%
  mutate(
    Dataset_label = paste(
      Dataset,
      " (S=",
      S_total,
      ", I=",
      I_total,
      ", R=",
      R_total,
      ")",
      sep = ""
    )
  ) %>% select(-`NA`, -`R`)

colSums(SIR_table[, -1])

validation_summary_table_total$Dataset_label[
  validation_summary_table_total$Dataset_label ==
    'All datasets combined (S=NA, I=NA, R=NA)'
] <- "All datasets combined (S=1757, I=356, R=4610)"

validation_summary_table_total <- 
  validation_summary_table_total %>% 
  mutate(Dataset_label = gsub("EUSCAPE & EURECA\\*", "PPPPPPPP", Dataset_label)) %>%
  mutate(Dataset_label = gsub("EUSCAPE & EURECA", "A (EUCAST, Various)", Dataset_label)) %>%
  mutate(Dataset_label = gsub("PPPPPPPP", "EUSCAPE & EURECA (Kvv, EUCAST)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("India GHRU", "B (EUCAST, V)", Dataset_label)) %>% 
  mutate(Dataset_label = gsub("AGAR GnSOP", "C (EUCAST, V/P)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("Oxfordshire\\*", "XXXXXXX", Dataset_label)) %>% 
  mutate(Dataset_label = gsub("Oxfordshire", "D (EUCAST, P)", Dataset_label)) %>% 
  mutate(Dataset_label = gsub("XXXXXXX", "Oxfordshire (Kvv, EUCAST)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("HURYC & CIBERINFEC", "E (EUCAST, S/M)", Dataset_label)) %>% 
  mutate(Dataset_label = gsub("HULP", "F (EUCAST, MW/E)", Dataset_label)) %>% 
  mutate(Dataset_label = gsub("JARBS-GNR", "G (CLSI, MW)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("MBIRA\\*", "YYYYYYYY", Dataset_label)) %>% 
  mutate(Dataset_label = gsub("MBIRA", "H (CLSI, V/E)", Dataset_label)) %>%
  mutate(Dataset_label = gsub("YYYYYYYY", "MBIRA (Kvv, CLSI)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("Controlling Superbugs study & Victorian CPE program", "I (CLSI, V/S)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("Queen Elizabeth Central Hospital", "J* (EUCAST/BSAC)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("SmaouiTunisia\\*", "K* (EUCAST)", Dataset_label)) %>% 
  
  mutate(Dataset_label = gsub("Addis Ababa University\\*", "L* (Kvv, CLSI)", Dataset_label)) %>%
  
  mutate(Dataset_label = gsub("University of Zurich", "M* (EUCAST)", Dataset_label)) %>%
  
  mutate(Dataset_label = gsub("CHAMPS", "N* (CLSI)", Dataset_label))

validation_summary_table_total$Dataset_label <- factor(
  as.character(validation_summary_table_total$Dataset_label),
  levels = c(
    "A (EUCAST, Various) (S=291, I=155, R=1687)", 
    "B (EUCAST, V) (S=84, I=11, R=441)",
    "C (EUCAST, V/P) (S=168, I=62, R=255)", 
    "D (EUCAST, P) (S=242, I=16, R=37)", 
    "E (EUCAST, S/M) (S=20, I=7, R=146)", 
    "F (EUCAST, MW/E) (S=11, I=1, R=136)", 
    "G (CLSI, MW) (S=222, I=163, R=449)",
    "H (CLSI, V/E) (S=109, I=14, R=128)", 
    "I (CLSI, V/S) (S=18, I=41, R=153)", 
    "J* (EUCAST/BSAC) (S=196, I=130, R=411)", 
    "K* (EUCAST) (S=62, I=28, R=146)", 
    "L* (Kvv, CLSI) (S=37, I=5, R=28)",
    "M* (EUCAST) (S=11, I=8, R=53)", 
    "N* (CLSI) (S=18, I=1, R=21)", 
    "IZSAM (S=7, I=3, R=273)",  
    "IRCCS Ospedale San Raffaele (S=9, I=2, R=187)", 
    "MBIRA (Kvv, CLSI) (S=85, I=0, R=5)",  
    "Oxfordshire (Kvv, EUCAST) (S=64, I=2, R=0)", 
    "SCHARKI (S=51, I=3, R=2)",
    "EUSCAPE & EURECA (Kvv, EUCAST) (S=35, I=3, R=3)", 
    "PachecoBrazil (S=10, I=0, R=28)", 
    "HarrisMERINO (S=7, I=8, R=21)", 
    "Overall median (datasets A-N)", 
    "All datasets combined (S=1757, I=356, R=4610)"
  )
)

validation_summary_table_total_exclude <- 
  validation_summary_table_total %>%
  filter(S_total > 10 | is.na(S_total)) %>%
  filter(R_total > 10 | is.na(R_total)) %>%
  filter(N_total > 25 | is.na(N_total))

validation_summary_table_median <-
  validation_summary_table_total_exclude %>% 
  group_by(`Evaluation metric`) %>% 
  summarise(metric_value = median(as.numeric(metric_value), na.rm = TRUE)) %>%
  mutate(Dataset = "Overall median (datasets A-N)") %>%
  mutate(var = "") %>%
  mutate(ub = "") %>%
  mutate(lb = "") %>%
  mutate(Country = "") %>%
  mutate(Prediction = "") %>%
  mutate(S_total = "") %>%
  mutate(I_total = "") %>%
  mutate(IR_total = "") %>%
  mutate(R_total = "") %>%
  mutate(N_total = "") %>%
  mutate(N = "") %>%
  mutate(Dataset_label = "Overall median (datasets A-N)")

validation_summary_table_total_exclude <-
  rbind(
    validation_summary_table_total_exclude,
    validation_summary_table_median
  )

validation_summary_table_total_exclude$lb[
  validation_summary_table_total_exclude$lb < 0
] <- 0

validation_summary_table_total_exclude <- 
  validation_summary_table_total_exclude %>%
  mutate(
    `Evaluation metric` = recode(
      `Evaluation metric`,
      "ME" = "Major errors"
    )
  ) %>%
  mutate(
    `Evaluation metric` = recode(
      `Evaluation metric`,
      "VME" = "Very major errors"
    )
  )

validation_summary_table_long_ci$`Evaluation metric` <- factor(
  validation_summary_table_long_ci$`Evaluation metric`,
  levels = c(
    "Categorical agreement",
    "Major errors",
    "Very major errors",
    "Minor errors"
  )
)



validation_summary_table_plot <- validation_summary_table_total_exclude %>% 
  ggplot(aes(as.numeric(metric_value), Dataset_label)) +
  
  geom_point(alpha = 0.5, col = "#9f98e9", size = 2) +
  
  geom_linerange(
    aes(
      xmin = as.numeric(lb) * 100,
      xmax = as.numeric(ub) * 100
    ),
    col = "#887feb",
    position = pd,
    size = 0.5
  ) +
  
  scale_y_discrete(limits = rev) +
  scale_x_continuous(breaks = seq(0, 100, 20)) +
  
  facet_grid(
    . ~ `Evaluation metric`,
    scales = "free_x",
    space = "free_x"
  ) +
  
  labs(y = "", x = "%") +
  
  theme_bw() +
  theme(
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  
  geom_vline(
    data = filter(
      validation_summary_table_total_exclude,
      `Evaluation metric` == "Categorical agreement"
    ),
    aes(xintercept = 90),
    colour = "#fb9472"
  ) +
  
  geom_vline(
    data = filter(
      validation_summary_table_total_exclude,
      `Evaluation metric` == "Major errors"
    ),
    aes(xintercept = 3),
    colour = "#fb9472"
  ) +
  
  geom_vline(
    data = filter(
      validation_summary_table_total_exclude,
      `Evaluation metric` == "Very major errors"
    ),
    aes(xintercept = 1.5),
    colour = "#fb9472"
  ) +
  
    geom_vline(
    data = filter(
      validation_summary_table_total_exclude,
      `Evaluation metric` == "Minor errors"
    ),
    aes(xintercept = 10),
    colour = "#fb9472"
  ) +

  ggh4x::force_panelsizes(
  cols = c(
    `Categorical agreement` = 1,
    `Major error` = 1,
    `VME` = 1,
    `Minor errors` = 1
  )
)


ggsave(width=10, height=4, file="../figs/Fig7_extValidation_summary.pdf")
ggsave(width=10, height=4, file="../figs/Fig7_extValidation_summary.png")
```


# TableS11 - eval metrics 
```{r, TableS11}

all_datasets_metrics <-
  validation_summary_table %>%

  mutate(`Categorical_agreement_conf_int_lower_bound`=paste0(round(as.numeric(cat_lb)*100,2),"%")) %>%
  mutate(`Categorical_agreement_conf_int_upper_bound`=paste0(round(as.numeric(cat_ub)*100,2),"%")) %>%
  mutate(`ME_conf_int_lower_bound`=paste0(round(as.numeric(me_lb)*100,2),"%")) %>%
  mutate(`ME_conf_int_upper_bound`=paste0(round(as.numeric(me_ub)*100,2),"%")) %>%
  mutate(`VME_conf_int_lower_bound`=paste0(round(as.numeric(vme_lb)*100,2),"%")) %>%
  mutate(`VME_conf_int_upper_bound`=paste0(round(as.numeric(vme_ub)*100,2),"%")) %>%
  mutate(`Sensitivity_conf_int_lower_bound`=paste0(round(as.numeric(sens_lb)*100,2),"%")) %>%
  mutate(`Sensitivity_conf_int_upper_bound`=paste0(round(as.numeric(sens_ub)*100,2),"%")) %>%
  mutate(`Specificity_conf_int_lower_bound`=paste0(round(as.numeric(spec_lb)*100,2),"%")) %>%
  mutate(`Specificity_conf_int_upper_bound`=paste0(round(as.numeric(spec_ub)*100,2),"%")) %>%
  mutate(`Minor_error_conf_int_lower_bound`=paste0(round(as.numeric(Minerror_lb)*100,2),"%")) %>%
  mutate(`Minor_error_conf_int_upper_bound`=paste0(round(as.numeric(Minerror_ub)*100,2),"%"))

all_datasets_metrics <- full_join(all_datasets_metrics, SIR_table) %>% select (-c(IR_total, N_total)) %>%
  rename(Susceptible=S_total)%>%
  rename(Intermediate=I_total)%>%
  rename(Resistant=R_total)

all_datasets_metrics <- all_datasets_metrics[, c("Dataset", "N","Susceptible", "Intermediate", "Resistant", "Prediction", "%R", "Categorical agreement", "Categorical_agreement_conf_int_lower_bound", "Categorical_agreement_conf_int_upper_bound", "ME", "ME_conf_int_lower_bound", "ME_conf_int_upper_bound", "VME", "VME_conf_int_lower_bound", "VME_conf_int_upper_bound", "Sensitivity", "Sensitivity_conf_int_lower_bound", "Sensitivity_conf_int_upper_bound", "Specificity", "Specificity_conf_int_lower_bound", "Specificity_conf_int_upper_bound", "Minor errors", "Minor_error_conf_int_lower_bound", "Minor_error_conf_int_upper_bound")]


all_datasets_metrics<- all_datasets_metrics %>% 
  mutate(Dataset = (gsub("EUSCAPE & EURECA\\*", "PPPPPPPP", Dataset)))%>%
  mutate(Dataset = (gsub("EUSCAPE & EURECA", "A (EUCAST)", Dataset)))%>%
  mutate(Dataset = (gsub("PPPPPPPP", "EUSCAPE & EURECA (Kvv, EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("India GHRU", "B (EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("AGAR GnSOP", "C (EUCAST)", Dataset)))%>% 
    mutate(Dataset = (gsub("Oxfordshire\\*", "XXXXXXX", Dataset)))%>% 
  mutate(Dataset = (gsub("Oxfordshire", "D (EUCAST)", Dataset)))%>% 
      mutate(Dataset = (gsub("XXXXXXX", "Oxfordshire (Kvv, EUCAST)", Dataset)))%>% 
    mutate(Dataset = (gsub("HURYC & CIBERINFEC", "E (EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("HULP", "F (EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("JARBS-GNR", "G (CLSI)", Dataset)))%>% 
      mutate(Dataset = (gsub("MBIRA\\*", "YYYYYYYY", Dataset)))%>% 
  mutate(Dataset = (gsub("MBIRA", "H (CLSI)", Dataset)))%>%
        mutate(Dataset = (gsub("YYYYYYYY", "MBIRA (Kvv, CLSI)", Dataset)))%>% 
  mutate(Dataset = (gsub("Controlling Superbugs study & Victorian CPE program", "I (CLSI)", Dataset)))%>% 
   mutate(Dataset = (gsub("Queen Elizabeth Central Hospital", "J* (EUCAST/BSAC)", Dataset)))%>% 
  mutate(Dataset = (gsub("SmaouiTunisia\\*", "K* (EUCAST)", Dataset)))%>% 
    mutate(Dataset = (gsub("Addis Ababa University\\*", "L* (Kvv, CLSI)", Dataset)))%>%
  mutate(Dataset = (gsub("University of Zurich", "M* (EUCAST)", Dataset)))%>%
  mutate(Dataset = (gsub("CHAMPS", "N* (CLSI)", Dataset)))

all_datasets_metrics <- all_datasets_metrics %>% 
  mutate_all(funs(str_replace(., "NA%", "-")))
#foundation for table S11 - some manual edits made to the table
write_tsv(all_datasets_metrics, "../tables/TableSX_extValidation_evalMetrics.tsv", na="-")
write_csv(all_datasets_metrics, "../tables/TableSX_extValidation_evalMetrics.csv", na="-")

```

#Table S11 - genotype profiles vs. PPVs and median MICs

# MIC prediction 

```{r, mic_prediction}
discrete_quantile <- function(x, p) {

  if (length(x)>5){
    sort(x)[length(x)*p]
  } else return(0)}

validation_geno <- read_delim("../tables/TableS5_genotypes_validation.tsv")
validation_pheno <- read_delim("../tables/TableS6_phenotype_metadata_validation.tsv")

validation <- full_join(validation_geno, validation_pheno)

#MIC median - with I excluded
MIC_pred_excludeI <- validation%>%
  filter(Laboratory.Typing.Method != "Disk diffusion") %>% 
  filter(Resistance.phenotype!="I") %>%
  group_by(Dataset,cipro_prediction_group) %>%
  summarise(median=median(Measurement),  
            lower25=discrete_quantile(as.numeric(Measurement), p=0.25), 
            upper25=discrete_quantile(as.numeric(Measurement), p=0.75)) %>% mutate(IQR_excludeI=paste0(median, " [", lower25, "-", upper25, "]"))
 
MIC_pred_excludeI[grep("0-0", MIC_pred_excludeI$IQR_excludeI), "IQR_excludeI"] <- "-"

MIC_pred_excludeI$lower25[MIC_pred_excludeI$lower25==0] <- "-"
MIC_pred_excludeI$lower25[MIC_pred_excludeI$upper25==0] <- "-"

MIC_pred_excludeI <- MIC_pred_excludeI %>% rename(Group=cipro_prediction_group) %>%
    select (-c(median,lower25,upper25))

MIC_pred_excludeI_pooled <- validation%>%
  filter(Laboratory.Typing.Method != "Disk diffusion") %>% 
  filter(Resistance.phenotype!="I") %>%
  group_by(cipro_prediction_group) %>%
  summarise(median=median(Measurement),  
            lower25=discrete_quantile(as.numeric(Measurement), p=0.25), 
            upper25=discrete_quantile(as.numeric(Measurement), p=0.75)) %>% mutate(IQR_excludeI=paste0(median, " [", lower25, "-", upper25, "]")) %>%
  mutate(Dataset=c("Pooled", "Pooled","Pooled", "Pooled","Pooled", "Pooled","Pooled", "Pooled","Pooled", "Pooled"))

MIC_pred_excludeI_pooled[grep("0-0", MIC_pred_excludeI_pooled$IQR_excludeI), "IQR_excludeI"] <- "-"

MIC_pred_excludeI_pooled$lower25[MIC_pred_excludeI_pooled$lower25==0] <- "-"
MIC_pred_excludeI_pooled$lower25[MIC_pred_excludeI_pooled$upper25==0] <- "-"

MIC_pred_excludeI_pooled <- MIC_pred_excludeI_pooled %>% rename(Group=cipro_prediction_group) %>%
    select (-c(median,lower25,upper25))


MIC_pred_excludeI <- rbind(MIC_pred_excludeI, MIC_pred_excludeI_pooled)



#MIC median - all included
MIC_pred <- validation%>%
  filter(Laboratory.Typing.Method != "Disk diffusion") %>% 
  group_by(Dataset,cipro_prediction_group) %>%
  summarise(median=median(Measurement),  
            lower25=discrete_quantile(as.numeric(Measurement), p=0.25), 
            upper25=discrete_quantile(as.numeric(Measurement), p=0.75)) %>% mutate(IQR=paste0(median, " [", lower25, "-", upper25, "]"))
 
MIC_pred[grep("0-0", MIC_pred$IQR), "IQR"] <- "-"

MIC_pred$lower25[MIC_pred$lower25==0] <- "-"
MIC_pred$lower25[MIC_pred$upper25==0] <- "-"

MIC_pred <- MIC_pred %>% rename(Group=cipro_prediction_group) %>%
    select (-c(median,lower25,upper25)) 

MIC_pred_SIR_pooled <- validation%>%
  filter(Laboratory.Typing.Method != "Disk diffusion") %>% 
  group_by(cipro_prediction_group) %>%
  summarise(median=median(Measurement),  
            lower25=discrete_quantile(as.numeric(Measurement), p=0.25), 
            upper25=discrete_quantile(as.numeric(Measurement), p=0.75)) %>% mutate(IQR=paste0(median, " [", lower25, "-", upper25, "]")) %>%
  mutate(Dataset=c("Pooled", "Pooled","Pooled", "Pooled","Pooled", "Pooled","Pooled", "Pooled","Pooled", "Pooled"))

MIC_pred_SIR_pooled[grep("0-0", MIC_pred_SIR_pooled$IQR), "IQR"] <- "-"

MIC_pred_SIR_pooled$lower25[MIC_pred_SIR_pooled$lower25==0] <- "-"
MIC_pred_SIR_pooled$lower25[MIC_pred_SIR_pooled$upper25==0] <- "-"

MIC_pred_SIR_pooled <- MIC_pred_SIR_pooled %>% rename(Group=cipro_prediction_group) %>%
    select (-c(median,lower25,upper25))

MIC_pred <- rbind(MIC_pred, MIC_pred_SIR_pooled)

```

# Table S12 - external validation by genotype profile
```{r}

#genotype profiles
pooled_dataset_genotype <- 
bind_rows(lst(MBIRA_Kpn_genotype,
          MBIRA_Kvv_genotype, 
          HowdenGorrie_genotype,
          AGAR_genotype,
          Milan_genotype,
          HailuEthiopia_genotype,
          KayamaJapan_genotype,
          FeaseyMalawi_genotype, 
          EUSCAPE_EURECA_Kpn_genotype,
          EUSCAPE_EURECA_Kvv_genotype,
          EgliSwitzerland_genotype,
          HRYC_Spain_genotype,
          IndiaGHRU_genotype,
          NeumannGermany_genotype,
          Whitelaw_CHAMPS_genotype,
          Oxfordshire_Kpn_genotype,
          Oxfordshire_Kvv_genotype,
          IZSAM_Italy_genotype, 
          LaPaz_Lazaro_genotype,
          HarrisMERINO_genotype,
          PachecoBrazil_genotype, 
          SmaouiTunisia_genotype, 
          pooled_genotype), .id ="Dataset") %>%
  mutate(Dataset = (gsub("_genotype", "", Dataset))) %>%
  mutate(N=S+I+R) %>%
  mutate(RexcludeI_PPV = paste0(R,"/",(S+R), " (", round(R/(S+R)*100, 2), "%)")) %>%
  mutate(R_PPV = paste0(R,"/",(S+I+R), " (", round(R/(S+I+R)*100, 2), "%)")) %>%
  mutate(NWT_PPV = paste0((R+I),"/",(S+I+R), " (", round((R+I)/(S+I+R)*100, 2), "%)")) %>%

  select (-c(S,I,R,N))


pooled_dataset_genotype$RexcludeI_PPV[pooled_dataset_genotype$RexcludeI_PPV == '0/0 (NaN%)'] <- '-'
pooled_dataset_genotype$R_PPV[pooled_dataset_genotype$R_PPV == '0/0 (NaN%)'] <- '-'
pooled_dataset_genotype$NWT_PPV[pooled_dataset_genotype$NWT_PPV == '0/0 (NaN%)'] <- '-'



pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'AGAR'] <- 'AGAR GnSOP'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'EUSCAPE_EURECA_Kpn'] <- 'EUSCAPE & EURECA'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'EUSCAPE_EURECA_Kvv'] <- 'EUSCAPE & EURECA*'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'EgliSwitzerland'] <- 'University of Zurich'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'FeaseyMalawi'] <- 'Queen Elizabeth Central Hospital'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'HRYC_Spain'] <- 'HURYC & CIBERINFEC'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'HailuEthiopia'] <- 'Addis Ababa University*'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'HowdenGorrie'] <- 'Controlling Superbugs study & Victorian CPE program'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'IZSAM_Italy'] <- 'IZSAM'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'IndiaGHRU'] <- 'India GHRU'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'KayamaJapan'] <- 'JARBS-GNR'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'LaPaz_Lazaro'] <- 'HULP'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'LaPaz_Lazaro'] <- 'HULP'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'MBIRA_Kpn'] <- 'MBIRA'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'MBIRA_Kvv'] <- 'MBIRA*'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'Milan'] <- 'IRCCS Ospedale San Raffaele'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'NeumannGermany'] <- 'SCHARKI'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'Oxfordshire_Kpn'] <- 'Oxfordshire'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'Oxfordshire_Kvv'] <- 'Oxfordshire*'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'SmaouiTunisia'] <- 'SmaouiTunisia'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'Whitelaw_CHAMPS'] <- 'CHAMPS'
pooled_dataset_genotype$Dataset[pooled_dataset_genotype$Dataset == 'pooled'] <- 'Pooled'

geno_MIC <- left_join(pooled_dataset_genotype, MIC_pred, by=c("Dataset","Group"))
geno_MIC <- left_join(geno_MIC, MIC_pred_excludeI, by=c("Dataset","Group"))

geno_MIC<- geno_MIC %>% 
  mutate(Dataset = (gsub("EUSCAPE & EURECA\\*", "PPPPPPPP", Dataset)))%>%
  mutate(Dataset = (gsub("EUSCAPE & EURECA", "A (EUCAST)", Dataset)))%>%
  mutate(Dataset = (gsub("PPPPPPPP", "EUSCAPE & EURECA (Kvv, EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("India GHRU", "B (EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("AGAR GnSOP", "C (EUCAST)", Dataset)))%>% 
    mutate(Dataset = (gsub("Oxfordshire\\*", "XXXXXXX", Dataset)))%>% 
  mutate(Dataset = (gsub("Oxfordshire", "D (EUCAST)", Dataset)))%>% 
      mutate(Dataset = (gsub("XXXXXXX", "Oxfordshire (Kvv, EUCAST)", Dataset)))%>% 
    mutate(Dataset = (gsub("HURYC & CIBERINFEC", "E (EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("HULP", "F (EUCAST)", Dataset)))%>% 
  mutate(Dataset = (gsub("JARBS-GNR", "G (CLSI)", Dataset)))%>% 
      mutate(Dataset = (gsub("MBIRA\\*", "YYYYYYYY", Dataset)))%>% 
  mutate(Dataset = (gsub("MBIRA", "H (CLSI)", Dataset)))%>%
        mutate(Dataset = (gsub("YYYYYYYY", "MBIRA (Kvv, CLSI)", Dataset)))%>% 
  mutate(Dataset = (gsub("Controlling Superbugs study & Victorian CPE program", "I (CLSI)", Dataset)))%>% 
   mutate(Dataset = (gsub("Queen Elizabeth Central Hospital", "J* (EUCAST/BSAC)", Dataset)))%>% 
  mutate(Dataset = (gsub("SmaouiTunisia", "K* (EUCAST)", Dataset)))%>% 
    mutate(Dataset = (gsub("Addis Ababa University\\*", "L* (Kvv, CLSI)", Dataset)))%>%
  mutate(Dataset = (gsub("University of Zurich", "M* (EUCAST)", Dataset)))%>%
  mutate(Dataset = (gsub("CHAMPS", "N* (CLSI)", Dataset)))

geno_MIC <- geno_MIC %>%
  rename("MIC median [IQR] (exclude I)"=IQR_excludeI) %>%
  rename("MIC median [IQR]"=IQR)
write_tsv(geno_MIC, "../tables/TableS12_extValidation_genotypeProfiles.tsv", na="-")
```



#plot - showing R, NWT, R_excludeI
```{r, eval=FALSE}

validation_summary_table_combined<-rbind(validation_summary_table, validation_summary_table_excludeI)

validation_summary_table_combined_long<-gather(validation_summary_table_combined, `Evaluation metric`, value, `Categorical agreement`:`%R`)

validation_summary_table_combined_long<- validation_summary_table_combined_long %>% mutate(value = as.numeric(gsub("%", "", value))) %>%
  mutate(`Evaluation metric` = (gsub("%R", "%R/NWT", `Evaluation metric`)))

validation_summary_table_combined_long$`Evaluation metric` = factor(validation_summary_table_combined_long$`Evaluation metric`, levels=c("Categorical agreement",'Sensitivity','Specificity',"ME", 'VME', "%R/NWT"))

validation_summary_table_median<-
validation_summary_table_combined_long %>% group_by(Prediction, `Evaluation metric`) %>% 
  summarise(value = median(as.numeric(value), na.rm = T)) %>%
  mutate(Dataset="Overall median") %>%
  mutate(Country="") %>%
  mutate(N="")

validation_summary_table_total <-
rbind(validation_summary_table_combined_long, validation_summary_table_median)

validation_summary_table_total <- full_join(validation_summary_table_total, SIR_table) %>% 
  mutate(Dataset = (gsub("Pooled", "All datasets combined", Dataset)))%>%
  mutate(Dataset_label=paste(Dataset, " (S=", S_total, ", I=", I_total, ", R=",R_total, ")", sep = ""))

validation_summary_table_total$Dataset_label[validation_summary_table_total$Dataset_label == 'Overall median (S=NA, I=NA, R=NA)'] <- 'Overall median (datasets A-N)'

# Get totals to explicitly write in label
colSums(SIR_table[,-1])
validation_summary_table_total$Dataset_label[validation_summary_table_total$Dataset_label == 'All datasets combined (S=NA, I=NA, R=NA)'] <- "All datasets combined (S=1757, I=652, R=4610)"

validation_summary_table_total$Dataset_label <- factor(validation_summary_table_total$Dataset_label, levels = c("EUSCAPE & EURECA (S=291, I=155, R=1687)", "JARBS-GNR (S=222, I=163, R=449)", "Queen Elizabeth Central Hospital (S=196, I=130, R=411)", "India GHRU (S=84, I=11, R=441)", "AGAR GnSOP (S=168, I=62, R=255)", "Oxfordshire (S=242, I=16, R=37)", "IZSAM (S=7, I=3, R=273)", "MBIRA (S=113, I=14, R=129)","SmaouiTunisia (S=62, I=28, R=146)",  "Controlling Superbugs study & Victorian CPE program (S=18, I=41, R=153)", "IRCCS Ospedale San Raffaele (S=9, I=2, R=187)", "HURYC & CIBERINFEC (S=20, I=7, R=146)", "HULP (S=11, I=1, R=136)", "MBIRA* (S=86, I=0, R=5)", "University of Zurich (S=11, I=8, R=53)", "Addis Ababa University* (S=37, I=5, R=28)", "Oxfordshire* (S=64, I=2, R=0)", "SCHARKI (S=51, I=3, R=2)", "CHAMPS (S=18, I=1, R=21)", "Overall median (datasets A-N)","All datasets combined (S=1757, I=652, R=4610)"))

validation_summary_table_plot <-ggplot(validation_summary_table_total, aes(as.numeric(value),fct_inorder(as.factor(Dataset_label)))) +
  geom_point(aes(shape = `Prediction`, size=`Prediction`, colour=`Prediction`), alpha=0.5) +
  scale_shape_manual(values = c(R=15, NWT=18, R_excludeI=20)) +
  scale_size_manual(values = c(R=3, NWT=3, R_excludeI=3)) +
  scale_colour_manual(values=c(R="#bd1515", NWT="#2c15ae", R_excludeI="#339966")) +
  scale_y_discrete(limits=rev)+
  facet_grid(`Evaluation metric`~., scales = "free", space='free')+ 
  labs(y="", x = "%") + 
  theme_bw()+
  theme(legend.title=element_blank())

#ggsave(width=10, height=14, file="../figs/FigX_extValidation.pdf")

# Only include datasets with at least 10 S genomes and 10 R genomes, >25 genomes in total, <100% S/R genomes
validation_summary_table_total_exclude <- 
validation_summary_table_total %>%
  filter(S_total>10 | is.na(S_total)) %>%
  filter(R_total>10 | is.na(R_total)) %>%
  filter(N>25 | is.na(R_total)) 

validation_summary_table_plot_exclude <-ggplot(validation_summary_table_total_exclude, aes(as.numeric(value),fct_inorder(as.factor(Dataset_label)))) +
  geom_point(aes(shape = `Prediction`, size=`Prediction`, colour=`Prediction`), alpha=0.5) +
  scale_shape_manual(values = c(R=15, NWT=18, R_excludeI=20)) +
  scale_size_manual(values = c(R=3, NWT=3, R_excludeI=3)) +
  scale_colour_manual(values=c(R="#bd1515", NWT="#2c15ae", R_excludeI="#339966")) +
  scale_y_discrete(limits=rev)+
  facet_grid(`Evaluation metric`~., scales = "free", space='free')+ 
  labs(y="", x = "%") + 
  theme_bw()+
  theme(legend.title=element_blank())

#ggsave(width=10, height=14, file="../figs/FigX_extValidation_exclude.pdf")

```

# Figure - external validation
```{r}
HowdenGorrie_MIC <- readPNG("HowdenGorrie_ValidationOutput/MIC_PredRIS.png", native=TRUE)
MBIRA_MIC_Kpn <- readPNG("MBIRA_ValidationOutput/MIC_PredRIS_Kpn.png", native=TRUE)
MBIRA_MIC_Kvv <- readPNG("MBIRA_ValidationOutput/MIC_PredRIS_Kvv.png", native=TRUE)
AGAR_MIC <- readPNG("AGAR_ValidationOutput/MIC_PredRIS.png", native=TRUE)
Milan_MIC <- readPNG("MilanSanRafaelle_ValidationOutput/MIC_PredRIS.png", native=TRUE)
HailuEthiopia_DD <- readPNG("HailuEthiopia_ValidationOutput/DD_PredRIS.png", native=TRUE)
KayamaJapan_MIC <- readPNG("KayamaJapan_ValidationOutput/MIC_PredRIS.png", native=TRUE)
FeaseyMalawi_DD <- readPNG("FeaseyMalawi_ValidationOutput/DD_PredRIS.png", native=TRUE)
EUSCAPE_EURECA_MIC <- readPNG("EURECA_EUSCAPE_ValidationOutput/MIC_PredRIS_Kpn.png", native=TRUE)


EgliSwitzerland_DD <- readPNG("EgliSwitzerland_2025_ValidationOutput/DD_PredRIS.png", native=TRUE)
HRYC_Spain_MIC <- readPNG("HRYC_Spain_ValidationOutput/MIC_PredRIS.png", native=TRUE)
IndiaGHRU_MIC <- readPNG("IndiaGHRU_2025_ValidationOutput/MIC_PredRIS.png", native=TRUE)
NeumannGermany_MIC <- readPNG("NeumannGermany_ValidationOutput/MIC_PredRIS.png", native=TRUE)

Whitelaw_CHAMPS_DD <- readPNG("Whitelaw_Stellenbosch_ValidationOutput/DD_PredRIS.png", native=TRUE)

Oxfordshire_Kpn_MIC <- readPNG("Oxfordshire_ValidationOutput/MIC_PredRIS_Kpn.png", native=TRUE)
Oxfordshire_Kvv_MIC <- readPNG("Oxfordshire_ValidationOutput/MIC_PredRIS_Kvv.png", native=TRUE)
IZSAM_Italy_MIC <- readPNG("IZSAM_Italy_ValidationOutput/MIC_PredRIS.png", native=TRUE)
LaPaz_Lazaro_MIC <- readPNG("IZSAM_Italy_ValidationOutput/MIC_PredRIS.png", native=TRUE)

HarrisMERINO_MIC <- readPNG("HarrisMERINO_ValidationOutput/MIC_PredRIS.png", native=TRUE)

PachecoBrazil_MIC <- readPNG("PachecoBrazil_ValidationOutput/MIC_PredRIS.png", native=TRUE)

TunisiaSmaoui_DD <- readPNG("SmaouiTunisia_ValidationOutput/DD_PredRIS.png", native=TRUE)


patch <- (wrap_elements(EUSCAPE_EURECA_MIC) +
            ggtitle("A) EUSCAPE & EURECA MIC (n=2133)")  +
            wrap_elements(KayamaJapan_MIC) +
            ggtitle("B) JARBS-GNR Japan MIC (n=834)")+
            wrap_elements(IndiaGHRU_MIC) +
            ggtitle("C) India GHRU MIC (n=536)")+
            wrap_elements(AGAR_MIC) + 
            ggtitle("D) AGAR GnSOP MIC (n=485)") +
            wrap_elements(Oxfordshire_Kpn_MIC) + 
            ggtitle("E) Oxfordshire MIC (n=295)") + 
            wrap_elements(IZSAM_Italy_MIC) +
            ggtitle("F) IZSAM MIC (n=263)")+            
            wrap_elements(MBIRA_MIC_Kpn) + 
            ggtitle("G) MBIRA MIC (n=263)")+ 
            wrap_elements(HowdenGorrie_MIC) +
            ggtitle("H) Controlling Superbugs study & \nVictorian CPE program MIC (n=212)") +
            wrap_elements(Milan_MIC) +
            ggtitle("I) IRCCS Ospedale San Raffaele MIC (n=198)")+
            wrap_elements(HRYC_Spain_MIC) +
            ggtitle("J) HURYC & CIBERINFEC MIC (n=173)")+
            wrap_elements(LaPaz_Lazaro_MIC) +
            ggtitle("K) HULP MIC (n=148)")+
            wrap_elements(MBIRA_MIC_Kvv) + 
            ggtitle("L) MBIRA* MIC (n=92)")+
            wrap_elements(Oxfordshire_Kvv_MIC) +
            ggtitle("M) Oxfordshire* MIC (n=66)")+
            wrap_elements(NeumannGermany_MIC) +
            ggtitle("N) SCHARKI MIC (n=56)")+  
            wrap_elements(PachecoBrazil_MIC) +
            ggtitle("O) Universidade Federal da \nBahia MIC (n=38)")+
            wrap_elements(HarrisMERINO_MIC) +
            ggtitle("P) MERINO Trial MIC (n=36)")+            
            
            wrap_elements(FeaseyMalawi_DD) +
            ggtitle("Q) Queen Elizabeth Central Hospital \n(Malawi) DD (n=737)") +
            wrap_elements(TunisiaSmaoui_DD) +
            ggtitle("R) Tunisia Smaoui DD (n=236)")+            
            wrap_elements(EgliSwitzerland_DD) +
            ggtitle("S) University of Zurich DD (n=72)")+
            wrap_elements(HailuEthiopia_DD) +
            ggtitle("T) Addis Ababa University* \n(Ethiopia) DD (n=70)")+
            wrap_elements(Whitelaw_CHAMPS_DD) +
            ggtitle("U) CHAMPS DD (n=40)")
          )+
  plot_layout(ncol = 3)

#ggsave(height=13, width=13, file="../figs/FigX_extValidation.png")
#ggsave(height=13, width=13, file="../figs/FigX_extValidation.pdf")

```




# Figure S11 - external validation (supplementary - distribution vs. genotype profile and phenotype vs. genotype profile)
```{r}
HowdenGorrie_dist <- readPNG("HowdenGorrie_ValidationOutput/MIC_PredGroup.png", native=TRUE)
HowdenGorrie_profile <- readPNG("HowdenGorrie_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)

MBIRA_Kpn_dist <- readPNG("MBIRA_ValidationOutput/MIC_PredGroup_Kpn.png", native=TRUE)
MBIRA_Kpn_profile <- readPNG("MBIRA_ValidationOutput/StackedBar_PredGroup_SIR_Kpn.png", native=TRUE)

MBIRA_Kvv_dist <- readPNG("MBIRA_ValidationOutput/MIC_PredGroup_Kvv.png", native=TRUE)
MBIRA_Kvv_profile <- readPNG("MBIRA_ValidationOutput/StackedBar_PredGroup_SIR_Kvv.png", native=TRUE)

AGAR_dist <- readPNG("AGAR_ValidationOutput/MIC_PredGroup.png", native=TRUE)
AGAR_profile <- readPNG("AGAR_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)

Milan_dist <- readPNG("MilanSanRafaelle_ValidationOutput/MIC_PredGroup.png", native=TRUE)
Milan_profile <- readPNG("MilanSanRafaelle_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)

Hailu_dist <- readPNG("HailuEthiopia_ValidationOutput/DD_PredGroup.png", native=TRUE)
Hailu_profile <- readPNG("HailuEthiopia_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)

Kayama_dist <- readPNG("KayamaJapan_ValidationOutput/MIC_PredGroup.png", native=TRUE)
Kayama_profile <- readPNG("KayamaJapan_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)

Feasey_dist <- readPNG("FeaseyMalawi_ValidationOutput/DD_PredGroup.png", native=TRUE)
Feasey_profile <- readPNG("FeaseyMalawi_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)
         
EUSCAPE_EURECA_dist <- readPNG("EURECA_EUSCAPE_ValidationOutput/MIC_PredGroup_Kpn.png", native=TRUE)
EUSCAPE_EURECA_profile <- readPNG("EURECA_EUSCAPE_ValidationOutput/StackedBar_PredGroup_SIR_Kpn.png", native=TRUE)    

EUSCAPE_EURECA_Kvv_dist <- readPNG("EURECA_EUSCAPE_ValidationOutput/MIC_PredGroup_Kvv.png", native=TRUE)
EUSCAPE_EURECA_Kvv_profile <- readPNG("EURECA_EUSCAPE_ValidationOutput/StackedBar_PredGroup_SIR_Kvv.png", native=TRUE)    

EgliSwitzerland_dist <- readPNG("EgliSwitzerland_2025_ValidationOutput/DD_PredGroup.png", native=TRUE)
EgliSwitzerland_profile <- readPNG("EgliSwitzerland_2025_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)           

HRYC_Spain_dist <- readPNG("HRYC_Spain_ValidationOutput/MIC_PredGroup.png", native=TRUE)
HRYC_Spain_profile <- readPNG("HRYC_Spain_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)           
IndiaGHRU_dist <- readPNG("IndiaGHRU_2025_ValidationOutput/MIC_PredGroup.png", native=TRUE)
IndiaGHRU_profile <- readPNG("IndiaGHRU_2025_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)           
NeumannGermany_dist <- readPNG("NeumannGermany_ValidationOutput/MIC_PredGroup.png", native=TRUE)
NeumannGermany_profile <- readPNG("NeumannGermany_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)           

WhitelawCHAMPS_dist <- readPNG("Whitelaw_Stellenbosch_ValidationOutput/DD_PredGroup.png", native=TRUE)
WhitelawCHAMPS_profile <- readPNG("Whitelaw_Stellenbosch_ValidationOutput/StackedBar_PredGroup_SIR_DD.png", native=TRUE)    

Oxfordshire_Kpn_dist <- readPNG("Oxfordshire_ValidationOutput/MIC_PredGroup_Kpn.png", native=TRUE)
Oxfordshire_Kpn_profile <- readPNG("Oxfordshire_ValidationOutput/StackedBar_PredGroup_SIR_Kpn.png", native=TRUE)    

Oxfordshire_Kvv_dist <- readPNG("Oxfordshire_ValidationOutput/MIC_PredGroup_Kvv.png", native=TRUE)
Oxfordshire_Kvv_profile <- readPNG("Oxfordshire_ValidationOutput/StackedBar_PredGroup_SIR_Kvv.png", native=TRUE)    

IZSAM_Italy_dist <- readPNG("IZSAM_Italy_ValidationOutput/MIC_PredGroup.png", native=TRUE)
IZSAM_Italy_profile <- readPNG("IZSAM_Italy_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)    
LaPaz_Lazaro_dist <- readPNG("LaPaz_Lazaro_ValidationOutput/MIC_PredGroup.png", native=TRUE)
LaPaz_Lazaro_profile <- readPNG("LaPaz_Lazaro_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)    

Harris_MERINO_dist <- readPNG("HarrisMERINO_ValidationOutput/MIC_PredGroup.png", native=TRUE)
Harris_MERINO_profile <- readPNG("HarrisMERINO_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)   

Pacheco_Brazil_dist <- readPNG("PachecoBrazil_ValidationOutput/MIC_PredGroup.png", native=TRUE)
Pacheco_Brazil_profile <- readPNG("PachecoBrazil_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)   

TunisiaSmaoui_dist <- readPNG("SmaouiTunisia_ValidationOutput/DD_PredGroup.png", native=TRUE)
TunisiaSmaoui_profile <- readPNG("SmaouiTunisia_ValidationOutput/StackedBar_PredGroup_SIR.png", native=TRUE)   

  (wrap_elements(EUSCAPE_EURECA_dist) +ggtitle("1) A (EUCAST, Various) / \n EURECA & EUSCAPE (n=2133) - MIC distribution vs.\n genotype profile") + wrap_elements(EUSCAPE_EURECA_profile) + ggtitle("2) A (EUCAST, Various) / \n EURECA & EUSCAPE (n=2133) - observed phenotype vs.\n genotype profile"))/
  
    (wrap_elements(IndiaGHRU_dist) +ggtitle("3) B (EUCAST, V) / \nIndia GHRU (n=536) - MIC distribution vs.\n genotype profile") + wrap_elements(IndiaGHRU_profile) + ggtitle("4) B (EUCAST, V) / \nIndia GHRU (n=536) - observed phenotype vs.\n genotype profile")) 
  


ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_1.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_1.pdf")

  (wrap_elements(AGAR_dist) +ggtitle("5) C (EUCAST, V/P) / \n AGAR GnSOP (n=485) - MIC distribution vs.\n genotype profile") + wrap_elements(AGAR_profile) + ggtitle("6) C (EUCAST, V/P) / \n AGAR GnSOP (n=485) - observed phenotype vs.\n genotype profile")) / 
  (wrap_elements(Oxfordshire_Kpn_dist) +ggtitle("7) D (EUCAST, P) / \n Oxfordshire (n=295) - MIC distribution vs.\n genotype profile") + wrap_elements(Oxfordshire_Kpn_profile) + ggtitle("8) D (EUCAST, P) / \n Oxfordshire (n=295) - observed phenotype vs.\n genotype profile")) 
  

ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_2.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_2.pdf")

  (wrap_elements(HRYC_Spain_dist) +ggtitle("9) E (EUCAST, S/M) / \n HURYC & CIBERINFEC (n=173) - MIC distribution vs.\n genotype profile") + wrap_elements(HRYC_Spain_profile) + ggtitle("10) E (EUCAST, S/M) \ \n HURYC & CIBERINFEC (n=173) \n - observed phenotype vs. genotype profile"))  / 
      (wrap_elements(LaPaz_Lazaro_dist) +ggtitle("11) F (EUCAST MW/E) / \n HULP (n=148) - MIC distribution vs.\n genotype profile") + wrap_elements(LaPaz_Lazaro_profile) + ggtitle("12) F (EUCAST, MW/E) \ \n HULP (n=148) \n - observed phenotype vs. genotype profile")) 

ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_3.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_3.pdf")

(wrap_elements(Kayama_dist) +ggtitle("13) G (CLSI, MW) / \n JARBS-GNR Japan (n=834) - MIC distribution vs.\n genotype profile") + wrap_elements(Kayama_profile) + ggtitle("14)  G (CLSI, MW) / \n JARBS-GNR Japan (n=834) - observed phenotype vs.\n genotype profile")) /
  
(wrap_elements(MBIRA_Kpn_dist) +ggtitle("15) H (CLSI, V/E) / \n MBIRA (n=263) - MIC distribution vs.\n genotype profile") + wrap_elements(MBIRA_Kpn_profile) + ggtitle("16) H (CLSI, V/E) / \n MBIRA (n=263) - observed phenotype vs.\n genotype profile")) 


ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_4.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_4.pdf")




(wrap_elements(HowdenGorrie_dist) +ggtitle("17)  I (CLSI, V/E) / \n Controlling Superbugs study & \nVictorian CPE program (n=212) - MIC distribution vs.\n genotype profile") + wrap_elements(HowdenGorrie_profile) + ggtitle("18)  I (CLSI, V/E) / \n Controlling Superbugs study & \nVictorian CPE program (n=212) - observed phenotype vs.\n genotype profile"))  /
(wrap_elements(Feasey_dist) +ggtitle("19) J* (EUCAST/BSAC) / \n Queen Elizabeth Central Hospital Malawi (n=737) \n - DD distribution vs. genotype profile") + wrap_elements(Feasey_profile) + ggtitle("20) J* (EUCAST/BSAC) / \n Queen Elizabeth Central Hospital Malawi (n=737) \n - observed phenotype vs. genotype profile")) 
  
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_5.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_5.pdf")


(wrap_elements(TunisiaSmaoui_dist) +ggtitle("21) K* (EUCAST) / \n Tunisia Smaoui* (n=236) .\n - DD distribution vs. genotype profile") + wrap_elements(TunisiaSmaoui_profile) + ggtitle("22) K* (EUCAST) / \n Tunisia Smaoui* (n=236) \n - observed phenotype vs. genotype profile")) /

(wrap_elements(Hailu_dist) +ggtitle("23) L* (Kvv, CLSI) / \n Addis Ababa University* (n=70) .\n - DD distribution vs. genotype profile") + wrap_elements(Hailu_profile) + ggtitle("24) L* (Kvv, CLSI) / \n Addis Ababa University* (n=70) \n - observed phenotype vs. genotype profile")) 

ggsave(height=13, width=10, file="../figs/FigS11_extValidation_genoprofiles_6.png")
ggsave(height=13, width=10, file="../figs/FigS11_extValidation_genoprofiles_6.pdf")

    (wrap_elements(EgliSwitzerland_dist) +ggtitle("25) M* (EUCAST) / \n University of Zurich (n=72) - \n DD distribution vs. genotype profile") + wrap_elements(EgliSwitzerland_profile) + ggtitle("26) M* (EUCAST) / \n University of Zurich (n=72) \n - observed phenotype vs. genotype profile")) /

(wrap_elements(WhitelawCHAMPS_dist) +ggtitle("27) N* (CLSI) / \n CHAMPS (n=40) - DD distribution vs.\n genotype profile") + wrap_elements(WhitelawCHAMPS_profile) + ggtitle("28) N* (CLSI) / \n CHAMPS (n=40) - observed phenotype vs.\n genotype profile")) 

ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_7.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_7.pdf")
  
  (wrap_elements(IZSAM_Italy_dist) +ggtitle("29) IZSAM (n=283) - MIC distribution vs.\n genotype profile") + wrap_elements(IZSAM_Italy_profile) + ggtitle("30) IZSAM (n=283) - observed phenotype vs.\n genotype profile")) /

(wrap_elements(Milan_dist) +ggtitle("31) IRCCS Ospedale San Raffaele (n=198) -\n MIC distribution vs. genotype profile") + wrap_elements(Milan_profile) + ggtitle("32) IRCCS Ospedale San Raffaele (n=198) \n - observed phenotype vs. genotype profile")) 

ggsave(height=13, width=10, file="../figs/FigS11_extValidation_genoprofiles_8.png")
ggsave(height=13, width=10, file="../figs/FigS11_extValidation_genoprofiles_8.pdf")

(wrap_elements(MBIRA_Kvv_dist) +ggtitle("33) MBIRA (Kvv, CLSI) (n=92) - MIC distribution vs.\n genotype profile") + wrap_elements(MBIRA_Kvv_profile) + ggtitle("34) MBIRA (Kvv, CLSI) (n=92) - observed phenotype vs.\n genotype profile")) /

(wrap_elements(Oxfordshire_Kvv_dist) +ggtitle("35) Oxfordshire (Kvv, EUCAST) (n=66) \n - MIC distribution vs. genotype profile") + wrap_elements(Oxfordshire_Kvv_profile) + ggtitle("36) Oxfordshire (Kvv, EUCAST) (n=66) \n - observed phenotype vs. genotype profile")) 

ggsave(height=13, width=10, file="../figs/FigS11_extValidation_genoprofiles_9.png")
ggsave(height=13, width=10, file="../figs/FigS11_extValidation_genoprofiles_9.pdf")

(wrap_elements(NeumannGermany_dist) +ggtitle("37) SCHARKI (n=56) - MIC distribution vs.\n genotype profile") + wrap_elements(NeumannGermany_profile) + ggtitle("38) SCHARKI (n=56) - observed phenotype vs.\n genotype profile")) /

(wrap_elements(EUSCAPE_EURECA_Kvv_dist) +ggtitle("39) EUSCAPE/EURECA Kvv (n=41) \n - MIC distribution vs. genotype profile") + wrap_elements(EUSCAPE_EURECA_Kvv_profile) + ggtitle("40) EUSCAPE/EURECA Kvv (n=41) \n - observed phenotype vs. genotype profile"))

ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_10.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_10.pdf")


(wrap_elements(Pacheco_Brazil_dist) +ggtitle("41) iberoAMRica study (n=38) \n - MIC distribution vs. genotype profile") + wrap_elements(Pacheco_Brazil_profile) + ggtitle("42) iberoAMRica study (n=38) \n - observed phenotype vs. genotype profile")) /
  
  (wrap_elements(Harris_MERINO_dist) +ggtitle("43) MERINO trial (n=36) \n - DD distribution vs. genotype profile") + wrap_elements(Harris_MERINO_dist) + ggtitle("44) MERINO trial (n=36) \n - observed phenotype vs. genotype profile")) 



ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_11.png")
ggsave(height=13, width=11, file="../figs/FigS11_extValidation_genoprofiles_11.pdf")



```






# numbers for summary

```{r load_data_summary, eval=FALSE}

validation_phenotypes <- read_delim("../tables/TableS6_phenotype_metadata_validation.tsv")
validation_genotypes <- read_delim("../tables/TableS5_genotypes_validation.tsv")

validation_phenotypes_summary_table <- validation_phenotypes %>% group_by(Dataset, Dataset_alias, `Assembly method`, Laboratory.Typing.Method, Testing.standard, Laboratory.Typing.Platform) %>% count()

#Create foundation for the table
#write_tsv(validation_phenotypes_summary_table, "../tables/TableS3_validation_summary.tsv", na="-")

```

```{r summary_genotypes_phenotypes, eval=FALSE}

# Have to add EUSCAPE + SAN RAF data 
# EUSCAPE = 2133 Kpn
# IRCCS Ospedale San Raffaele = 198 Kpn 
validation_genotypes %>% count(species)

# EUSCAPE = 2133 MIC
# IRCCS Ospedale San Raffaele = 198 MIC
validation_phenotypes %>% count(Laboratory.Typing.Method)

```

